### 2009-08-31 AffineLieAlgebras/InformalNotes/ModularInvariantsAndBranching.tex
Informal notes are moved into separate folder, backup files are now ignored in git. Start of the work on note on modular invariance in conformal field theory.

\documentclass[a4paper,12pt]{article}
\usepackage[unicode,verbose]{hyperref}
\usepackage{amsmath,amssymb,amsthm} \usepackage{pb-diagram}
\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage[russian]{babel}
\usepackage{cmap}
\usepackage{graphicx}
\pagestyle{plain}
\theoremstyle{definition} \newtheorem{Def}{Definition}
\begin{document}
\tableofcontents

\bibliography{CFTNotes}{}
\bibliographystyle{utphys}

\end{document}