### 2010-09-30 GeneralOverview/IntegrableSystems.tex
some changes in integrable systems

\documentclass[a4paper,12pt]{article}
\usepackage[unicode,verbose]{hyperref}
\usepackage{amsmath,amssymb,amsthm} \usepackage{pb-diagram}
\usepackage{ucs}
\usepackage{cmap}
\usepackage{graphicx}
\pagestyle{plain}
\theoremstyle{definition} \newtheorem{Def}{Definition}
\newenvironment{comment}
{\par\noindent{\bf Comment}\\}
{\\\hfill$\scriptstyle\blacksquare$\par}

\newtheorem{statement}{Statement}
\theoremstyle{definition} \newtheorem{Def}{Definition}
\newcommand{\go}{\overset{\circ }{\frak{g}}}
\newcommand{\ao}{\overset{\circ }{\frak{a}}}
\newcommand{\co}[1]{\overset{\circ }{#1}}

\begin{document}

\title{Integrable systems and representation theory}
\begin{abstract}
  We review basic notions of quantum integrable systems and then discuss the applications of the representation theory of finite-dimensional and affine Lie algebras.
\end{abstract}
\begin{document}

\bibliography{program}{}
\bibliographystyle{utphys}
\end{document}