### 2011-11-10 AffineLieAlgebras/Thesis/concl.tex
Thesis work

\conclusion

\chapter{Заключение}
\label{cha:conclusion1}

%%
%% End of file
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: