### 2011-12-13 AffineLieAlgebras/Thesis/5.tex
Translation of Affine.m description started

\chapter{Вычислительные методы для теории представлений аффинных алгебр Ли}
\label{cha:computational-methods}

В данной главе мы описываем вычислительный пакет  {\bf Affine.m}, разработанный нами на основе идей и методов, изложенных в главах \ref{cha:affine-lie-algebras}, \ref{cha:BGG}. Этот пакет для популярной системы компьютерной алгебры {\it Mathematica} может использоваться для вычислений в теории представлений конечномерных и аффинных алгебр Ли. Реализованные в пакете алгоритмы основываются на свойствах весов и симметрии Вейля. Основные проблемы, которые может решать данный пакет -- это вычисление кратностей весов в неприводимых модулях и модулях Верма, построение правил ветвления и функций ветвления, а также разложение тензорных произведений. Такие задачи важны с точки зрения физических приложений (см. главы \ref{cha:CFT},\ref{sec:SLE}) и в данной главе мы приводим дополнительные примеры.

Вычислительные методы в теории представлений имеют долгую историю \cite{belinfante1989survey}, существует большое количество программ и пакетов для вычислений, связанных с алгебрами Ли \cite{simplie}, \cite{vanleeuwen1994lsp}, \cite{stembridge1995mps,coxweyl}, \cite{fischbacher2002ilp}, \cite{Fuchs:1996dd}.

Большинство популярных программ \cite{simplie}, \cite{vanleeuwen1994lsp}, \cite{fischbacher2002ilp}, \cite{coxweyl} создано для изучения теории представлений простых конечномерных алгебр Ли. Здесь основные вычислительные задачи это:
\begin{enumerate}
\item Построение корневой системы, определяющей свойства алгебры Ли, в том числе коммутационные соотношения.
\item Перечисление элементов группы Вейля, необходимое ввиду симметрии корневой системы и характеров представлений относительно группы Вейля.
\item Вычисление кратностей весов, коэффициентов ветвления и слияния
\end{enumerate}
Существуют различные алгоритмы для решения этих задач \cite{moody1982fast}, \cite{stembridge2001computational}, \cite{belinfante1989survey}, \cite{casselman1994machine}.
Третья задача наиболее сложна с точки зрения вычислений. Существует два рекуррентных алгоритма, основывающихся на формуле Вейля для характеров и формуле Фрейденталя для кратностей. В данной главе мы их опишем.

Infinite-dimensional Lie algebras also have growing number of applications in physics for example in conformal field theory and in a study of  integrable systems. But infinite-dimensional algebras are much harder to investigate and the number of available computer programs is much smaller in this case.

Affine Lie algebras \cite{kac1990idl} constitute important and tractable class of infinite-dimensional Lie algebras. They are constructed as central extensions of loop algebras of (semi-simple) finite-dimensional Lie algebras and appear naturally in a study of Wess-Zumino-Witten and coset models in conformal field theory \cite{Walton:1999xc}, \cite{difrancesco1997cft}, \cite{Goddard198588}, \cite{Dunbar:1992gh}.

The structure of affine Lie algebras allows to adapt for them computational algorithms created for finite-dimensional Lie algebras  \cite{Fuchs:1996dd}, \cite{gannon2001algorithms}, \cite{kass1990ala}. The book \cite{kass1990ala} with the tables of multiplicities and other computed characteristics of affine Lie algebras and representations was published in 1990. But we are not aware of software packages for popular computer algebra systems which can be used to extend these results.
We address this issue and present {\bf Affine.m} -- {\it Mathematica} package for computations in representation theory of affine and finite-dimensional Lie algebras.  We describe the features and limitations of the package in the present paper.  We also provide representation-theoretical background of implemented algorithms and present examples of computations relevant to physical applications.

The paper starts with an overview of Lie algebras and their representation theory (Sec. \ref{sec:theor-backgr}). Then we describe datastructures of {\bf Affine.m} used to present different objects related to Lie algebras and representations (Sec. \ref{sec:core-datastructures}) and discuss implemented algorithms (Sec. \ref{sec:comp-algor}). Next section consists of physically interesting examples (Sec. \ref{sec:examples}). The paper is concluded with the discussion of possible extensions and refinements (Sec. \ref{sec:conclusion}).

\section{Theoretical background}
\label{sec:theor-backgr}

In this section we remind necessary definitions and present formulae used in computations.

%% Define external border of irreducible representation here

We can use the Weyl character formula to obtain recurrent relations for weight multiplicities -- important tools for calculations \cite{il2010folded,kulish4sfa}.

For irreducible highest-weight modules recurrent relations have the following form
\begin{equation}
\label{eq:14}
m_{\xi }=-\sum_{w\in W\setminus e}\epsilon (w)m_{\xi
-\left( w(\rho )-\rho \right) }+\sum_{w\in W}\epsilon
(w)\delta _{\left( w(\mu +\rho )-\rho \right) ,\xi }.
\end{equation}
Formulae for Verma and generalized Verma modules differ only in the second term on the right-hand side. In case of Verma module it is just $\delta_{\xi,\mu}$. For generalized Verma module the summation in the second term on the right-hand side of \eqref{eq:14} is over the Weyl subgroup generated by reflections corresponding to roots $\{\alpha_{I}\}$.

Other recurrent formula can be obtained from a study of Casimir
elements action on irreducible highest weight modules
\cite{humphreys1997introduction}:
\begin{equation}
  \label{eq:15}
  m_{\lambda}=\frac{2}{(\mu+\rho)^{2}-(\lambda+\rho)^{2}}\sum_{\alpha\in \Delta^{+}}\sum_{k\geq 1} (\lambda+k\alpha,\alpha)m_{\lambda+k\alpha}.
\end{equation}
It is called the Freudenthal multiplicity formula.
Note that it is applicable only to irreducible modules.

We discuss the use of formulae \eqref{eq:14} and \eqref{eq:15} for computations in section \ref{sec:comp-algor}.

Now consider an algebra $\gf$ and a reductive subalgebra
$\af\subset \gf$. Simple roots $\beta_{i}$ of the subalgebra $\af$
can be presented as linear combinations of $\gf$-algebra roots
$\alpha_{j}$: $\beta_{i}=\sum_{j=1,\dots,r_{\gf}}k_{j}
\alpha_{j},\ j=1,\dots,r_{\af}$.

Each irreducible $\gf$-module is also an $\af$-module, although
$L^{\mu}_{\gf}$ is in general not irreducible as $\af$-module. 
It can be
decomposed into a direct sum of irreducible $\af$-modules:
\begin{equation}
  \label{eq:16}
  L^{\mu}_{\gf}=\bigoplus_{\nu}b^{\mu}_{\nu}L^{\nu}_{\af}
\end{equation}
Coefficients in such a decomposition are called branching
coefficients.

It is possible to calculate branching coefficients by constructing and successively  subtracting the submodules $L^{\nu}_{\af}$. 
This traditional approach has serious limitations especially in case of affine Lie algebras. We discuss them in the end of section \ref{sec:comp-algor}.

Now we describe an alternative approach which is based on recurrent properties of branching coefficients. 
But before we proceed to these recurrent relations we need several additional definitions.

For a subalgebra $\af\subset \gf$ we introduce the subalgebra
$\afb$. Consider the root subspace $\hf_{\perp \af}^{\ast }$
orthogonal to $\hf_{\af}$,
\begin{equation*}
\hf_{\perp \af}^{\ast }:=\left\{ \eta \in \hf^{\ast }|\forall
h\in \hf_{\af};\eta \left( h\right) =0\right\} ,
\end{equation*}
and the roots (correspondingly -- positive roots) of $\gf$ orthogonal
to roots of $\af$,
\begin{eqnarray}
\Delta _{\afb} &:&=\left\{ \beta \in \Delta _{\gf}|\forall
h\in \hf_{\af};\beta \left( h\right) =0\right\} ,
\label{delta a ort} \\
\Delta _{\afb}^{+} &:&=\left\{ \beta ^{+}\in \Delta _{\gf%
}^{+}|\forall h\in \hf_{\af};\beta ^{+}\left( h\right) =0\right\} .
\notag
\end{eqnarray}
Let $W_{\afb}$ be a subgroup of $W$ generated by
reflections $w_{\beta }$ with the roots $\beta \in \Delta _{\af_{\perp
}}^{+}$. The subsystem $\Delta _{\afb}$ determines a 
subalgebra $\afb$ with the Cartan subalgebra $\hf_{\af%
_{\perp }}$.

The Cartan subalgebra $\frak{h}$ can be decomposed in the following way:  $\frak{h}=\frak{\frak{h}_{\af}}\oplus
\frak{h}_{\afb}\oplus \frak{h}_{\perp }$

We also introduce the notations
\begin{eqnarray}
\widetilde{\frak{a}_{\perp }} :=\frak{a}_{\perp }\oplus \frak{h}_{\perp }
\qquad
\widetilde{\frak{a}} :=\frak{a}\oplus \frak{h}_{\perp }.
\end{eqnarray}

For $\af$ and $\afb$ we consider the
corresponding Weyl vectors, $\rho _{\af}$ and $\rho _{\af_{\perp
}} $ and compose the so called ''defects'' $\mathcal{D}_{\af}$ and $\mathcal{%
D}_{\afb}$ of the injection:
\begin{equation}
\mathcal{D}_{\af}:=\rho _{\af}-\pi _{\af}\rho , \qquad
\mathcal{D}_{\afb}:=\rho _{\afb}-\pi _{\af%
_{\perp }}\rho .  \label{defect-ort}
\end{equation}

%%!!! Define P^+
For $\mu \in P^{+}$ consider the linked weights $\left\{ \left(
w(\mu +\rho )-\rho \right) |w\in W\right\} $ and their projections
to
$h_{\afb}^{\ast }$ additionally shifted by the defect $-%
\mathcal{D}_{\afb}$:
\begin{equation*}
\mu _{\afb}\left( w\right) :=\pi _{\afb}\left[
w(\mu +\rho )-\rho \right] -\mathcal{D}_{\afb},\quad w\in W.
\end{equation*}
Among the weights $\left\{ \mu _{\af_{\perp
}}\left( w\right) |w\in W\right\} $ one can always choose those located in
the fundamental chamber $\overline{C_{\afb}}$. Let $U$ be the
set of representatives $u$ for the classes $W/W_{\afb}$ such
that

\begin{equation}
U:=\left\{ u\in W|\quad \mu _{\afb}\left( u\right) \in
\overline{C_{\afb}}\right\} \quad .  \label{U-def}
\end{equation}
Thus we can form the subsets:
\begin{equation}
\mu _{\widetilde{\mathfrak{a}}}\left( u\right) :=\pi _{\widetilde{%
\mathfrak{a}}}\left[ u(\mu +\rho )-\rho \right] +\mathcal{D}_{\af%
_{\perp }},\quad u\in U,  \label{mu-a}
\end{equation}
and
\begin{equation}
\mu _{\afb}\left( u\right) :=\pi _{\afb}\left[
u(\mu +\rho )-\rho \right] -\mathcal{D}_{\afb},\quad u\in U.
\label{mu-a-tilda}
\end{equation}

Notice that the subalgebra $\mathfrak{a}_{\bot}$ is regular by definition
since it is built on a subset of roots of the algebra $\mathfrak{g}$.

Denote by  $k_{\xi }^{\left( \mu \right) }$ signed branching coefficients. If $\xi\in \bar C_{\af}$ is in main Weyl chamber $k_{\xi}^{(\mu)}=b^{(\mu)}_{\xi}$ otherwise $k_{\xi}^{(\mu)}=\epsilon(w) b^{(\mu)}_{w (\xi+\rho_{\af})-\rho_{\af}}$ where $w\in W_{\af}$ is such that $w (\xi+\rho_{\af})-\rho_{\af}\in \bar C_{\af}$.

Now we can use the Weyl character formula to write a
recurrent relation \cite{2010arXiv1007.0318L} for signed branching
coefficients $k_{\xi }^{\left( \mu \right) }$ corresponding to an
injection $\af\hookrightarrow \gf$:
\begin{equation}
\begin{array}{c}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
\sum_{u\in U}\epsilon (u)\;\dim \left( L_{\afb}^{\mu _{\af%
_{\perp }}\left( u\right) }\right) \delta _{\xi -\gamma _{0},\pi _{%
\widetilde{\af}}(u(\mu +\rho )-\rho )}+\right.  \\
\left. +\sum_{\gamma \in \Gamma _{\widetilde{\af}\rightarrow \gf%
}}s\left( \gamma +\gamma _{0}\right) k_{\xi +\gamma }^{\left( \mu \right)
}\right) .
\end{array}
\label{recurrent-rel}
\end{equation}
The recursion is governed by the set $\Gamma _{\af\rightarrow \gf}$ called the injection fan. The latter is defined by the
carrier set $\left\{ \xi \right\} _{\af\rightarrow \gf}$ for the
coefficient function $s(\xi )$
\begin{equation*}
\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}:=\left\{
\xi \in P_{\widetilde{\af}}|s(\xi )\neq 0\right\}
\end{equation*}
appearing in the expansion
\begin{equation}
\prod_{\alpha \in \Delta ^{+}\setminus \Delta _{\bot }^{+}}\left( 1-e^{-\pi
_{\widetilde{\af}}\alpha }\right) ^{\mathrm{mult}(\alpha )-\mathrm{mult}%
_{\af}(\pi _{\widetilde{\af}}\alpha )}=-\sum_{\gamma \in P_{%
\widetilde{\af}}}s(\gamma )e^{-\gamma };\quad
\end{equation}
The weights in $\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}$ are to be shifted by $\gamma _{0}$ -- the lowest vector in $\left\{ \xi
\right\} $ -- and the zero element is to be eliminated:
\begin{equation}
\Gamma _{\af\rightarrow \gf}=\left\{ \xi -\gamma
_{0}|\xi \in \left\{ \xi \right\} \right\} \setminus \left\{ 0\right\} .
\end{equation}
Formula (\ref{eq:14}) is a particular case of recurrent relation for branching coefficients (\ref{recurrent-rel}) in the case of Cartan subalgebra $\af=\hfg$.

If the root system of $\afb$ is generated by some subset of $\gf$
simple roots $\alpha_{1},\dots,\alpha_{r}$ then the recurrent
relation (\ref{recurrent-rel}) is connected with the generalized
Bernstein-Bernstein-Gelfand resolution for parabolic Verma modules
\cite{2011arXiv1102.1702L}.

Another particular case of this formula is connected with tensor
product decompositions. Consider the tensor product of two
irreducible $\gf$-modules $L^{\mu}\otimes L^{\nu}$. It is also a
$\gf$-module but not irreducible in general. So
\begin{equation}
  \label{eq:19}
  L^{\mu}\otimes L^{\nu}=\bigoplus_{\gamma} f^{\mu\nu}_{\gamma}L^{\gamma}
\end{equation}
The coefficients $f^{\mu\nu}_{\gamma}$ are called fusion
coefficients. The problem of computation of fusion coefficients is
equivalent to branching problem for the diagonal subalgebra
$\gf\subset \gf\oplus \gf$ (see \cite{LyakhovskyPostnova2011}). So our implementation of recurrent algorithm can be used to decompose tensor products (See Section \ref{sec:tens-prod-decomp}).

In the case of affine Lie algebras $\gf, \af$  the multiplicities
$m_{\nu}$ and the corresponding branching coefficients $b_{\nu}$
can be regarded as the coefficients in power series decomposition
of string and branching functions correspondingly:
\begin{align}
  \label{eq:21}
  &\sigma_{\nu}(q)=\sum_{n=0}^{\infty} m_{\nu-n\delta} q^n, \quad \nu=\sum_j c_j \omega_j,\quad c_j\geq 0\\
  & b_{\nu}(q)=\sum_{n=0}^{\infty} b_{\nu-n\delta} q^n,\quad  \nu=\sum_j c_j \omega_j, \quad c_j\geq 0
\end{align}
String and branching functions have  modular and analytic properties which are important for conformal field theory, especially in  coset models and the study of CFT on higher genus surfaces \cite{kac1988modular}, \cite{difrancesco1997cft}, \cite{Walton:1999xc}, \cite{walton1989conformal}.

{\bf Affine.m}  calculates weight multiplicities and branching coefficients for affine Lie algebras up to some finite grade. We present examples of computations in Sections \ref{sec:string-funct-affine}, \ref{sec:branch-funct-coset}. Now we proceed to the description of datastructures and algorithms implemented in {\bf Affine.m}.

\section{Core datastructures}
\label{sec:core-datastructures} Having introduced necessary
mathematical objects, problems and relations we now describe the
related datastructures of {\bf Affine.m}. Although {\it
Mathematica} is untyped language it is possible to create
structured objects and do type checks with patterns
\cite{shifrinmathematica}, \cite{maeder2000computer}.
\subsection{Weights}
\label{sec:weights}

Weights are represented  by two data-structures: \lstinline{finiteWeight} for finite-dimensional Lie algebras and \lstinline{affineWeight}  for affine.

Internally the finite weight is a \lstinline{List} with the
\lstinline{Head} \lstinline{finiteWeight}, its components are the
coordinates of the weight in the orthogonal Bourbaki  basis
\cite{bourbaki2002lie}.

Affine weight is an extension of a finite weight by supplying it
with level and grade coordinates. There is a set of functions
defined for finite and affine weights. The complete list can be
found in  online help of the package. The most important
are definitions of addition, multiplication by a number and a
scalar product (bilinear form) for weights. These definitions
allow to use traditional notation when working with {\bf
Affine.m}:
\begin{lstlisting}
  w=makeFiniteWeight[{1,0,3}];
  v=makeFiniteWeight[{3,2,1}];
  2*w+v==makeFiniteWeight[{5,2,7}]
  w.v==6
\end{lstlisting}

The use of orthogonal basis in the internal structure of weights allows us to work with weights without complete specification of root system which is useful for study of branching, since roots of the subalgebra can be specified by hand.

\subsection{Root systems}
\label{sec:root-systems}

To specify an algebra of finite or affine type it is enough to fix
its root system. Root systems are represented by two datatypes
\lstinline{finiteRootSystem} and \lstinline{affineRootSystem}. The
latter is an extension of the former. We offer several different
constructors for these datastructures. It is possible to specify
the set of simple roots explicitly, for example to study the
subalgebra $B_2\subset B_4$ we can use the definition
\begin{lstlisting}
  b2b4=makeFiniteRootSystem[ { {1,-1,0,0}, {0,1,0,0} } ]
\end{lstlisting}
There are constructors for root systems of simple finite-dimensional Lie algebras:
\begin{lstlisting}
  b2=makeSimpleRootSystem[B,2]
\end{lstlisting}
We use typographic features of {\it Mathematica} frontend to offer mathematical notation for simple Lie algebras:

\begin{lstlisting}[mathescape=true]
  $B_2$ == makeFiniteRootSystem[ { {1, -1}, {0, 1} }]
\end{lstlisting}

Non-twisted affine root systems can be created as affine extensions of finte root systems, e.g.
\begin{lstlisting}
  b2affine = makeAffineExtension[b2]
\end{lstlisting}
In notebook interface this can be written simply as $\hat{B}_2$.

Semisimple Lie algebras can be created as sums of simple:
\begin{lstlisting}[mathescape=true]
  $A_1\oplus A_1$ == finiteRootSystem[2, 2, {finiteWeight[2, {1, 0}], finiteWeight[2, {0, 1}]}]
\end{lstlisting}
%% The problem is here

Predicate \lstinline{rootSystemQ} checks if the object is root systems of finite or affine type.

List of simple roots is the property of root system so it is accessed as \lstinline{rs[simpleRoots]}.

We have implemented several functions to get major properties of root systems. Weyl vector is given by the function \lstinline{rho[rs_?rootSystemQ]}:
\begin{lstlisting}[label=list:1]
  In[1]  =  rho[b2]
  Out[1]  =  finiteWeight[2, {3/2, 1/2}]
\end{lstlisting}
Positive roots can be constructed with \lstinline{positiveRoots[rs_?rootSystemQ]}. For an affine Lie algebra this and related functions return the list up to fixed grade. This grade limit is set through the value of \lstinline{rs[gradeLimit]} which is equal to 10 by default. List of roots (up to \lstinline{gradeLimit}) is returned by \lstinline{roots[rs]}. Cartan matrix and fundamental weights are calculated by the functions \lstinline{cartanMatrix} and \lstinline{fundamentalWeights} correspondingly.

It is possible to specify a weight of a Lie algebra by its Dynkin labels
\begin{lstlisting}
  weight[b2][1,2] == makeFiniteWeight[{2, 1}]
\end{lstlisting}
The function \lstinline{dynkinLabels[rs_?rootSystemQ][wg_?weightQ]} returns Dynkin labels of weight \lstinline{wg} in the root system \lstinline{rs}.

Elements of Weyl group are specified by the set of indices of reflections, so the element  $w=s_{1}s_{2}s_{1}$ of the Weyl group of algebra $B_{2}$ is constructed with \lstinline{weylGroupElement[b2][1,2,1]}. Then it can be applied to the weights:
\begin{lstlisting}
  w = weylGroupElement[b2][1,2,1];
  w @ makeFiniteWeight[{1,0}] == makeFiniteWeight[{-1,0}]
\end{lstlisting}

Computation of lexicographically minimal form \cite{casselman1994machine,casselman1995automata} for Weyl group elements can be conveniently implemented through the use of pattern-matching in {\it Mathematica}. In \cite{KallenShortlex} rewrite rules for simple finite dimensional and affine Lie algebras are presented as {\it Mathematica} patterns. Our presentation of Weyl group elements is compatible with the code of \cite{KallenShortlex}
\begin{lstlisting}[mathescape=true]
  In[1]  = $<<$A3reduce;
           reduce[s[1,2,1,2,1,3,2,1,1]]
  Out[1] = s[2, 3, 2]

  In[2]  = (weylGroupElement[$A_{3}$] @@ reduce[s[1,2,1,2,1,3,2,1,1]]) @ weight[$A_{3}$][-1,-2,-1]
  Out[2] = finiteWeight[4, {-2, 2, 1, -1}]

  In[3]  = dynkinLabels[$A_{3}$][Out[2]]
  Out[3] = {-4, 1, 2}
\end{lstlisting}

\subsection{Formal elements}
\label{sec:formal-elements}

We represent formal characters of representations by special structure \lstinline{formalElement}. This structure is a hash-table implemented with \lstinline{DownValues}. The keys are weights at the exponents and the values are the  corresponding multiplicities. \lstinline[mathescape=true]!makeFormalElement[{$\gamma_{1},\dots,\gamma_{n}$},{$m_{1},\dots,m_{n}$}]! creates data-structure which represents the element $\sum_{i=1}^{n} m_{i} e^{\gamma_{i}}$ of the formal algebra $\mathcal{E}$. The operations in $\mathcal{E}$ are implemented for \lstinline{formalElement} data-type: formal elements can be added, multiplied by a number or by an exponent of a weight. There exists also a multiplication of formal elements but no division.
\begin{lstlisting}[mathescape=true]
  In[1]  = makeFormalElement[{makeFiniteWeight[{1,1}],makeFiniteWeight[{0,0}]},{1,2}] *
             (2 * Exp[makeFiniteWeight[{1,0}]] *
             makeFormalElement[{makeFiniteWeight[{1,1}],makeFiniteWeight[{0,0}]},{1,2}]);
  In[2]  = In[1][weights]
  Out[2] = {finiteWeight[2, {1, 0}], finiteWeight[2, {2, 1}], finiteWeight[2, {3, 2}]}

  In[3]  = In[1][multiplicities]
  Out[3] = {8, 8, 2}
\end{lstlisting}

\subsection{Modules}
\label{sec:modules}

{\bf Affine.m} can be used to study different kinds of modules, i.e. Verma modules, irreducible modules and parabolic Verma modules.  We need datastructure \lstinline{module} to represent generic module of Lie algebra $\gf$. Module properties can be deduced from its set of singular weights using Weyl character formulae \eqref{eq:11},\eqref{eq:12},\eqref{eq:18},\eqref{eq:13}. Set of singular weights can have Weyl symmetry. It can be a symmetry with respect to the Weyl group $W_{\gf}$ or with respect to some subalgebra $W_{\af}$ as in the case of parabolic Verma modules. Then it is possible to study only the main Weyl chamber $C_{\af}$. To use this symmetry generic constructor for \lstinline{module} datastructure accepts several parameters \lstinline{makeModule[rs_?rootSystemQ][singWeights_formalElement,subs_?rootSystemQ|emptyRootSystem[],limit:10}. Here \lstinline{rs} is the root system of the Lie algebra $\gf$,\lstinline{singWeights} is the set of singular weights,  \lstinline{subs} is the root system corresponding to the Weyl group $W_{\af}$ which is the (anti-)symmetry of the set of singular weights. Parameter \lstinline{limit} limits computation for infinite-dimensional modules such as Verma or parabolic Verma.
There are several specialized constructors for different types of highest weight modules:
\begin{lstlisting}[mathescape=true]
vm=makeVermaModule[$B_{2}$][{2,1}];
pm=makeParabolicVermaModule[$B_{2}$][weight[$B_{2}$][2,1],{1}];
im=makeIrreducibleModule[$B_{2}$][2,1];
GraphicsRow[textPlot/@{im,vm,pm}]
$\includegraphics[width=130mm]{figures/irrep-verma-pverma}$
\end{lstlisting}

As we already stated properties of a module are encoded by its singular element. Function \lstinline{singularElement[m_module]} returns singular element of a module as \lstinline{formalElement} datastructure. Character (up to \lstinline{limit} for (parabolic) Verma modules) is returned by the function \lstinline{character[m_module]}. Direct sum of modules is a module and we use natural notation
\begin{lstlisting}[mathescape=true]
im1=makeIrreducibleModule[$B_{2}$][weight[$B_{2}$][2,1]];
im2=makeIrreducibleModule[$B_{2}$][weight[$B_{2}$][1,2]];
textPlot[im1$\oplus$ im2]
$\includegraphics[width=60mm]{figures/irrep-sum}$
\end{lstlisting}

The tensor product is also implemented but only for finite-dimensional Lie algebras, since tensor product of affine Lie algebra modules leads to rich new structures \cite{kazhdan1994tensor3,kazhdan1993tensor1,kazhdan1993tensor2} which are out of the scope of the present paper.
\begin{lstlisting}[mathescape=true]
textPlot[makeIrreducibleModule[$A_{1}$][5]$\otimes$ makeIrreducibleModule[$A_{1}$][3]];
$\includegraphics[width=60mm]{figures/tensor-product}$
\end{lstlisting}
\section{Computational algorithms}
\label{sec:comp-algor}

As we have already stated in section \ref{sec:high-weight-modul} there exist two recurrent relations which can be used to calculate weight multiplicities in irreducible modules. Both algorithms proceed in the following way to calculate weight multiplicities:
\begin{enumerate}
\item Create the list of weights in the main Weyl chamber by subtracting all possible combinations of simple roots from the highest weight (e.g. for finite-dimensional algebra subtract $\alpha_{1}$ from $\mu$ while inside $\bar C$, then subtract $\alpha_{2}$ from all the weights already obtained etc).
\item Sort the list of weights by their product with the Weyl vector.
\item Use a recurrent formula. If the weight required for recurrent computation is outside the main chamber use the Weyl symmetry.
\end{enumerate}
The difference in the performance of algorithms is in the number of previous values required to get the multiplicity of weight under consideration. For  the Weyl formula based recurrent relation \eqref{eq:14} it is constant and equal to the number of elements in the Weyl group (if we are far from the boundary of representation diagram). When the Freudenthal formula \eqref{eq:15} is used number of previous values grows with the distance from the external border of representation. So the Freudenthal formula is faster if the weight is close to the border or the rank of the algebra and the size of Weyl group is large \cite{moody1982fast}.
Note that the Freudenthal formula is valid for irreducible modules only, so it can not be used to study (generalized) Verma modules.

We have made some experiments with our implementations of the Freudenthal formula and formula \eqref{eq:15} and have got Figure  \ref{fig:freudenthal-racah-times}, which depicts the dependence of the computation time on the number of weights in a module.

\begin{figure}[h]
  \noindent\centering{
    \includegraphics[width=80mm]{figures/timing}
  }
  \caption{Running time of algorithms based on the Freudenthal formula \eqref{eq:15} (dashed) and recurrent relation \eqref{eq:14} (solid) with the number of weights in $\bar C$ for calculation of multiplicities in representations of $B_{2}$.}
\label{fig:freudenthal-racah-times}

\end{figure}

In the calculation of branching coefficients the application of the Freudenthal formula requires a complete construction of formal characters of an algebra module and all the representations of a subalgebra. It is impractical when the rank of algebra and subalgebra are big, for example for maximal subalgebras.

Alternative algorithm which was presented in the paper \cite{2010arXiv1007.0318L}  contains the following steps:

\begin{enumerate}
\item  Construct the root system $\Delta _{\af}$ for the embedding $%
\af\rightarrow \frak{g}$.

\item  Select all the positive roots $\alpha \in \Delta ^{+}$ orthogonal
to  $\af$, i.e. form the set $\Delta_{\afb }^{+}$.

\item  Construct the set $\Gamma _{\af\rightarrow \frak{g}}$. Relation
 (\ref{eq:6}) defines the sign function
 $s(\gamma)$ and the set $\Phi_{\af\subset \frak{g}}$ where the lowest weight
 $\gamma_0$ is to be subtracted to get the fan:
 $\Gamma _{\af\rightarrow \frak{g}}=\left\{ \xi -\gamma _{0}|\xi \in \Phi _{%
\af\subset \frak{g}}\right\} \setminus \left\{ 0\right\}$.

\item  Construct the set $\widehat{\Psi ^{(\mu )}}=\left\{ w (\mu +\rho
)-\rho ;\;w \in W\right\} $ of singular weights for the $\frak{g}$%
-module $L^{(\mu )}$.

\item  Select the weights $\left\{ \mu _{\widetilde{\af_{\perp }}%
}\left( w\right) =\pi _{\widetilde{\af_{\perp }}}\left[ w(\mu +\rho
)-\rho \right] -\mathcal{D}_{\af_{\perp }}\in \overline{C_{\widetilde{%
\af_{\perp }}}}\right\} $. Since the set $\Delta_{\afb }^{+}$ is fixed
we can easily check wether the weight $\mu _{\widetilde{\af_{\perp }}%
}\left( w\right) $ belongs to the main Weyl chamber $\overline{C_{\widetilde{%
\af_{\perp }}}}$ (by computing its scalar product with the fundamental
weights of $\afb^{+}$).

\item  For the weights $\mu _{\widetilde{\af_{\perp }}}\left( w\right) $
calculate dimensions of the corresponding modules, $\mathrm{\dim }\left(
L_{\widetilde{\af_{\perp }}}^{\mu _{\widetilde{\af_{\perp }}%
}\left( u\right) }\right) $, using the Weyl dimension formula and construct
the singular element $\Psi ^{\left( \mu \right) }_{\left(  \af, \afb \right)}$.

\item  Calculate the anomalous branching coefficients using 
recurrent relation (\ref{recurrent-rel}) and select among them those
corresponding to the weights in the main Weyl
chamber $\overline{C_{\af}}$.
\end{enumerate}

We can speed up the algorithm by
one-time computation of the representatives of the conjugate classes $%
W/W_{\afb }$.

Consider the regular embedding $B_{2}\subset B_{4}$. In this case the fan consists of 24 elements. In order to decompose $B_{4}$ module we need to construct the subset of singular weights of the module which projects to the main Weyl chamber of the subalgebra $B_{2}$. The full set of singular weights consists of 384 elements. The required subset contains at most 48 elements. So the time for the construction of this required subset is negligible if the number of branching coefficients is greater than that. We may estimate the total number of required operations for the computation of branching coefficients as the product of the number of elements in the main Weyl chamber of a subalgebra with non-zero branching coefficients and the number of elements in the fan. In the case of a direct algorithm we need to compute the multiplicities for each module in the decomposition, so the number of operations grows faster than the square of the number of elements in the main Weyl chamber of a subalgebra with non-zero branching coefficients.

To further illustrate this performance issue we include Figure \ref{fig:branching} where we show the time required for computations of branching coefficients for $B_{3}\subset B_{4}$.

\begin{figure}[h]
  \noindent\centering{
   \includegraphics[width=100mm]{figures/branching-timing}
  }
  \caption{Running time of algorithms based on the Freudenthal formula \eqref{eq:15} (dashed) and recurrent relation \eqref{recurrent-rel} (solid) with the number of weights in $\bar C$ for calculation of branching coefficients for $B_{3}\subset B_{4}$.}
  \label{fig:branching}
\end{figure}

\section{Examples}
\label{sec:examples}
In this section we present some examples of computations available with {\bf Affine.m} with the code required to produce these results.

\subsection{Tensor product decompositon for finite-dimensional Lie algebras}
\label{sec:tens-prod-decomp}

Computation of fusion coefficients for the decomposition of tensor product of highest-weight modules to the direct sum of irreducible modules has numerous applications in physics. For example, we can consider the spin of a composite system such as an atom. Another interesting example is integrable spin chain consisting of $N$ particles with the spins living in some representation $L$ of a Lie algebra $\gf$ with a $\gf$-invariant Hamiltonian $H$, describing the nearest-neighbour spin-spin interaction. In order to solve such a system, i.e. to find eigenstates of the Hamiltonian, we need to decompose $L^{\otimes N}$ into the direct sum of irreducible $\gf$-modules of lower dimension and diagonalize the Hamiltonian on these modules.

For fundamental representations of simple Lie algebras it is sometimes possible to get an analytic result describing the dependence of decomposition coefficients on $N$ (See \cite{LyakhovskyPostnova2011}). Our code provides numerical values and can be used to check this analytic results.

Consider a tensor power of $B_{2}$ the first fundamental representation $\left(L^{[1,0]}\right)^{\otimes 4}$. Decomposition coefficients are just the branching
 coefficients for the tensor power module reduced to the diagonal subalgebra $B_{2}\subset B_{2}\oplus B_{2}\oplus B_{2}\oplus B_{2}$. So the following code calculates these coefficients:
\begin{lstlisting}[mathescape=true]
fm = makeIrreducibleModule[$B_{2}$][1, 0];
tp = ((fm$\otimes$ ]fm)$\otimes$ fm)$\otimes$]fm;
subs = makeFiniteRootSystem[
  {1/4*{1, -1, 1, -1, 1, -1, 1, -1}, 
   1/4*{0, 1, 0, 1, 0, 1, 0, 1}}];
bc = branching[tp, subs];
{bc[#], dynkinLabels[subs][#]} & /@ bc[weights]
\end{lstlisting}
It produces a list of highest weights and tensor product decomposition coefficients:
\begin{lstlisting}
{{1, {4, 0}}, {3, {2, 2}}, {0, {3, 0}}, 
{2, {0, 4}}, {3, {1, 2}}, {6, {2, 0}}, 
{6, {0, 2}}, {1, {1, 0}}, {3, {0, 0}}}]
\end{lstlisting}

Returning to the problem of spin chain Hamiltonian diagonalization we can see that instead of diagonalizing operator in space of dimension $625$ we can diagonalize operators in the spaces of dimensions $55, 81, 30, 35, 35, 14, 10, 5, 1$.

\subsection{Branching and parabolic Verma modules}
\label{sec:branch-parab-verma}

We illustrate the generalized BGG-resolution by the diagrams of $G_{2}$ parabolic Verma modules which appear in the decomposition of irreducible module $L^{[1,1]}_{G_{2}}$:
\begin{equation}
\mathrm{ch}\left( L^{\mu }\right) =\sum_{u\in U}\;e^{\mu _{\aft}\left(
u\right) }\epsilon (u)\mathrm{ch}M_{I}^{\mu _{\frak{a}_{\perp }}\left(
u\right) }.  \label{char-in-gen-verma-mod}
\end{equation}
The character of $L^{[1,1]}$ is presented in Figure \ref{branching-bgg}, characters of the generalized Verma modules in decomposition \eqref{char-in-gen-verma-mod} are shown in Figure \ref{g2-pverma}. Characters in the upper row appear in \eqref{char-in-gen-verma-mod} with positive sign and in the lower row -- with negative.

\begin{figure}[h]
  \noindent\centering{
    \includegraphics[width=80mm]{figures/G2-irrep}
  }
  \caption{Character of the irreducible $G_{2}$-module $L^{[1,1]}$}
  \label{branching-bgg}
\end{figure}
\begin{figure}[h]
  \noindent\centering{
    \includegraphics[width=150mm]{figures/G2-pverma}
  }
  \caption{Characters of $G_{2}$ generalized Verma modules appearing in the decomposition of $L^{[1,1]}$. Parabolic Verma modules in the upper row appear in the decomposition with a positive sign, in the lower row -- with a negative.}
  \label{g2-pverma}
\end{figure}

\subsection{String functions of affine Lie algebras and CFT models}
\label{sec:string-funct-affine}

String functions can be used to present formal characters of affine Lie algebra highest weight representation. They have interesting analytic and modular properties \cite{kac1990idl,kac1988modular,kac1984infinite}.

{\bf Affine.m} produces power series decomposition for string functions. Consider an affine Lie algebra $\hat{sl(3)}=\hat A_{2}$ and its highest weight module $L^{(1,0,0)}$. To get the string functions we can use the code:
\begin{lstlisting}[mathescape=true]
stringFunctions[$\hat A_2$,{1,1,2}]
{{0, 0, 4}, 
  $2 q + 10 q^2 + 40 q^3 + 133 q^4 + 398 q^5 + 1084 q^6 + 2760 q^7 + 6632 q^8 + 15214 q^9 + 33508 q^{10}$}, 
{{0, 3, 1}, 
  $2 q + 12 q^2 + 49 q^3 + 166 q^4 + 494 q^5 + 1340 q^6 + 3387 q^7 + 8086 q^8 + 18415 q^9 + 40302 q^{10}$}, 
{{1, 1, 2}, 
  $1 + 6 q + 27 q^2 + 96 q^3 + 298 q^4 + 836 q^5 + 2173 q^6 + 5310 q^7 + 12341 q^8 + 27486 q^9 + 59029 q^{10}$}, 
{{2, 2, 0}, 
  $1 + 8 q + 35 q^2 + 124 q^3 + 379 q^4 + 1052 q^5 + 2700 q^6 + 6536 q^7 + 15047 q^8 + 33248 q^9 + 70877 q^{10}$}, 
{{3, 0, 1}, 
  $2 + 12 q + 49 q^2 + 166 q^3 + 494 q^4 + 1340 q^5 + 3387 q^6 + 8086 q^7 + 18415 q^8 + 40302 q^9 + 85226 q^{10}$}
\end{lstlisting}
Similarly for the affine Lie algebra $\hat G_{2}$ we get
\begin{lstlisting}[mathescape=true]
stringFunctions[$\hat G_2$,{1,1,0}]
{{2, 0, 0}, 
  $1 + 8 q + 37 q^2 + 138 q^3 + 431 q^4 + 1227 q^5 + 3208 q^6 + 7901 q^7$}, 
{{0, 0, 1}, 
  $3 q + 18 q^2 + 73 q^3 + 247 q^4 + 736 q^5 + 2000 q^6 + 5070 q^7$}, 
{{1, 1, 0},
  $1 + 7 q + 32 q^2 + 117 q^3 + 370 q^4 + 1055 q^5 + 2780 q^6 + 6880 q^7$}, 
{{0, 2, 0}, 
  $3 q + 15 q^2 + 63 q^3 + 210 q^4 + 633 q^5 + 1725 q^6 + 4407 q^7$}
\end{lstlisting}

\subsection{Branching functions and coset models of conformal field theory}
\label{sec:branch-funct-coset}

It is believed that most rational models of CFT can be obtained from cosets $G/A$ corresponding to the embedding $\af\subset\gf$. These models can be studied as gauge theories \cite{Hwang:1994yr, hwang1993brst}.

Branching functions for an embedding $\af\subset\gf$ are the partition functions of CFT on the torus (see \cite{difrancesco1997cft}).

As a first example we show how to construct branching functions for the embedding $\hat A_{1}\to \hat B_{2}$ up to the tenth grade:
\begin{lstlisting}[mathescape=true]
branchingFunctions[$\hat B_{2}$,makeAffineExtension[makeFiniteRootSystem[{{1, 1}}]], {1, 1, 1}]

 {{3, 0}, 
  $2 + 14 q + 52 q^2 + 154 q^3 + 410 q^4 + 994 q^5 + 2248 q^6 + 4832 q^7 + 9934 q^8 + 19680 q^9 + 37802 q^{10}$},
 {{2, 1}, 
  $4 + 20 q + 72 q^2 + 220 q^3 + 584 q^4 + 1424 q^5 + 3248 q^6 + 7012 q^7 + 14488 q^8 + 28844 q^9 + 55616 q^{10}$},
 {{0, 3}, 
  $4 q + 20 q^2 + 68 q^3 + 200 q^4 + 516 q^5 + 1224 q^6 + 2736 q^7 + 5808 q^8 + 11820 q^9 + 23236 q^{10}$},
 {{1, 2}, 
  $2 + 14 q + 54 q^2 + 168 q^3 + 462 q^4 + 1148 q^5 + 2656 q^6 + 5812 q^7 + 12130 q^8 + 24358 q^9 + 47328 q^{10}$}
\end{lstlisting}

Another example demonstrates the computation of branching functions for the regular embedding $\hat B_{2}\subset \hat C_{3}$:
\begin{lstlisting}[mathescape=true]
sub=makeAffineExtension[parabolicSubalgebra[$C_{3}$][2,3]];
branchingFunctions[$\hat C_{3}$,sub, {2, 0, 0, 0}]

{{0, 1, 0}, 
  $2 q - 20 q^3 + 24 q^4 + 82 q^5 - 320 q^6 + 108 q^7$}, 
{{1, 0, 0}, 
  $1 - q - 8 q^2 + 19 q^3 + 16 q^4 - 156 q^5 + 205 q^6 + 640 q^7$}, 
{{0, 0, 1}, 
  $q - 5 q^3 + 7 q^5$}
\end{lstlisting}

\section{Conclusion}
\label{sec:conclusion}
We have presented the package {\bf Affine.m} for computations in representation theory of finite-dimensional and affine Lie algebras. It can be used to study Weyl symmetry, root systems, irreducible, Verma and parabolic Verma modules of finite-dimensional and affine Lie algebras.  In the present paper we have also discussed main ideas used for an implementation of the package and described most important notions of representation theory required to use {\bf Affine.m}.

We have demonstrated that recurrent approach based on the Weyl character formula is not only useful for calculations but also allows to establish connections with the (generalized) Bernstein-Bernstein-Gelfand resolution.

Also we have presented examples of computations with this package connected with problems of physics and mathematics.

In future versions of our software we are going to treat twisted affine Lie algebras, extended affine Lie algebras and provide more direct support for tensor product decompositions.

\section*{Acknowledgements}
\label{sec:acknowledgements}
I thank V.D. Lyakhovsky and O.Postnova for discussions and helpful comments.

The work is supported by the Chebyshev Laboratory
(Department of Mathematics and Mechanics, Saint-Petersburg State
University) under the grant 11.G34.31.0026 of the Government of the
Russian Federation.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%\appendix

\section{Software package}
\label{package}
The package can be freely downloaded from \url{http://github.com/naa/Affine}. To get the development code use the command
\begin{lstlisting}[language=bash]
 git clone git://github.com/naa/Affine.git
\end{lstlisting}

Contents of the package:
\begin{verbatim}
    Affine/                                root folder
      demo/                                  demonstrations
        demo.nb                                demo notebook
        paper.nb                               code for the paper
      doc/                                 documentation folder
        figures/                             figures in paper 
          timing.pdf                           diagram showing performance
          branching-timing.pdf                 ...  for branching coefficients  
          irrep-sum.pdf                        sum of B2 irreps
          irrep-verma-pverma.pdf               irrep, Verma, (p)Verma for B2
          G2-irrep.pdf                         irrep for G2
          G2-pverma.pdf                        parabolic Verma for G2
          tensor-product.pdf                   tensor product of A1-modules
        bibliography.bib                     bibliographic database
        paper.pdf                            present paper
        paper.tex                            paper source
        TODO.org                             list of issues
      src/                                 source folder
        affine.m                             main software package
      tests/                               unit tests folder
        tests.m                              unit tests
      README.markdown                      installation and usage notes
\end{verbatim}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%% References with bibTeX database:

  %   \section*{References}
  %   \label{sec:references}
  %   
  %   
  %   \bibliography{bibliography}
  %   \bibliographystyle{phcpc}
  %   %\bibliographystyle{model1-num-names}
  %   
  %   
  %   %% Authors are advised to submit their bibtex database files. They are
  %   %% requested to list a bibtex style file in the manuscript if they do
  %   %% not want to use elsarticle-num.bst.
  %   
  %   %% References without bibTeX database:
  %   
  %   % \begin{thebibliography}{00}
  %   
  %   %% \bibitem must have the following form:
  %   %%   \bibitem{key}...
  %   %%
  %   
  %   % \bibitem{}
  %   
  %   % \end{thebibliography}
  %   
  %   
  %

%%
%% End of file
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: