### 2012-03-17 AffineLieAlgebras/SLE/paper.tex
Some work on SLE paper

\documentclass[a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{multicol}
\usepackage{color}
\usepackage{graphicx}
\usepackage{hyperref}

\newtheorem{statement}{Statement}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}{Lemma}
\newtheorem{mynote}{Note}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newcommand{\go}{\stackrel{\circ }{\mathfrak{g}}}
\newcommand{\ao}{\stackrel{\circ }{\mathfrak{a}}}
\newcommand{\co}[1]{\stackrel{\circ }{#1}}
\newcommand{\pia}{\pi_{\mathfrak{a}}}
\newcommand{\piab}{\pi_{\mathfrak{a}_{\bot}}}
\newcommand{\gf}{\mathfrak{g}}
\newcommand{\af}{\mathfrak{a}}
\newcommand{\aft}{\widetilde{\mathfrak{a}}}
\newcommand{\afb}{\mathfrak{a}_{\bot}}
\newcommand{\hf}{\mathfrak{h}}
\newcommand{\hfb}{\mathfrak{h}_{\bot}}
\newcommand{\pf}{\mathfrak{p}}
\newcommand{\gfh}{\hat{\mathfrak{g}}}
\newcommand{\afh}{\hat{\mathfrak{a}}}
\newcommand{\bff}{\mathfrak{b}}
\newcommand{\hfg}{\hf_{\gf}}

\theoremstyle{definition} \newtheorem{Def}{Definition}
\newcommand{\adk}{a^{\dagger}_{\kappa}} \newcommand{\ak}{a_{\kappa}}
\def\bF{\mbox{$\overline{\cal F}$}} \def\F{\mbox{$\cal F$}}

\begin{document}

\title{SLE martingales in coset conformal field theory}
\author{Anton Nazarov}
\maketitle
\section{Introduction}
\label{sec:introduction}

Generalizing results of papers \cite{bettelheim2005stochastic,santachiara2008sle,alekseev2010sle} we
consider Schramm-Loewner evolution $SLE_{\kappa}$ with additional Brownian motion on coset space
$G/A$ characterised by parameter $\tau$ proposed in our paper \cite{2011arXiv1112.4354N}. We consider correlation functions in coset models of
conformal field theory which satisfy martingale conditions with respect to this stochastic process.
Since all minimal unitary models of conformal field theory can be obtained from coset construction
for certain choice of Lie group $G$ and subgroup $A$ \cite{difrancesco1997cft} these correlation functions correspond to SLE
observables appearing as the scaling limit of lattice observables. Trace of
Schramm-Loewner $\gamma_{t}$ evolution describes critical domain wall in lattice models.

 Consider upper half-plane $\mathbb{H}$ with the cut along a critical interface $\gamma_{t}$. We denote this slit domain by $\mathbb{H}_{t}=\mathbb{H}\setminus \gamma_{t}$. Conformal map from $\mathbb{H}_{t}$ to $\mathbb{H}$ is denoted by $g_{t}:\mathbb{H}_{t}\to \mathbb{H}$ and satisfies stochastic differential equation \cite{schramm2000scaling}:
\begin{equation}
\label{eq:19}
  \frac{\partial g_t(z)}{\partial t} = \frac{ 2}{g_t(z)-\sqrt{\kappa}\xi_{t}} ,
\end{equation}
where $\xi_{t}$ is the Brownian motion. The dynamic of the tip $z_{t}$ of critical curve $\gamma_{t}$ (tip of SLE trace) is given by the law $z_{t}=g_{t}^{-1}(\sqrt{\kappa}\xi_{t})$.

For us it is more convenient to use map $w_{t} (z)=g_{t}(z)-\sqrt{\kappa}\xi_{t}$, so the equation \eqref{eq:19} becomes
\begin{equation}
  \label{eq:20}
       d w _{t}= \frac{2dt}{w_{t} }-\sqrt{\kappa}\xi_{t}  
\end{equation}

It is natural to study the problem in upper half-plane  and consider correlation
functions of boundary conformal field theory with boundary condition changing operators sitting at
the tip of SLE trace and at the infinity.

So we are to consider $N$-point boundary correlation functions on the upper half-plane of the form 
\begin{equation}
  \mathcal{F}(\left\{z_{i},\bar z_{i}\right\}_{i=1}^{N})_{\mathbb{H}_{t}}=
\left<\varphi(z_{t}) \phi_{1}(z_{1},\bar z_{1}) \dots \phi_{n}(z_{n},\bar z_{n})
      \varphi(\infty)\right>,
    \label{eq:2}
\end{equation}
where $\phi_{i}$ are bulk primary fields and $\varphi(z_{t})$ is boundary condition changing operator. We assume conformal invariance of boundary conditions so boundary states satisfy Cardy conditions \cite{cardy1989boundary,cardy1986effect,cardy1984conformal}.

 We consider Schramm-Loewner evolution with additional Brownian motion on $G/A$-coset space \cite{2011arXiv1112.4354N}. Denote by $\mathcal{K}$ Killing form of Lie algebra $\gf$ corresponding to Lie group $G$ and by $t^{a}_{i}$ ($\tilde{t}^{b}_{i}$) the  generators of $\gf$-representation ($\af$-representation) corresponding to primary field $\phi_{i}$.
Primary fields in coset models are labeled by pairs $(\lambda,\eta)$ of weights $\gf$ and $\af$-representations. For each $\lambda$ possible values of $\eta$ are those which appear in decomposition of representation $L^{(\mu)}_{\gf}=\bigoplus_{\eta} H^{\lambda}_{\eta}\otimes L^{(\eta)}_{\af}$. Some pairs are equivalent, we discuss and use this equivalence relation later. 
We assume that boundary field $\varphi$ is also primary and labeled by weights $(\mu,\nu)$.

  We use the conformal map  $w(z):\mathbb{H}\setminus\gamma_{t}\to \mathbb{H}$ to rewrite expression \eqref{eq:2} in the whole upper half plane:

\begin{equation}
  \mathcal{F}(\left\{z_{i},\bar z_{i}\right\})_{\mathbb{H}_{t}}=\prod \left(\frac{\partial w(z_{i})}{\partial z_{i}}\right)^{h_{i}} 
  \prod \left(\frac{\partial \bar w(\bar z_{i})}{\partial \bar z_{i}}\right)^{\bar h_i}
  \mathcal{F}(\left\{w_{i}, \bar w_{i}\right\})_{\mathbb{H}}
  \label{eq:1}
\end{equation}
We can extend the definition of $\mathcal{F}$ to the whole plane by doubling the number of fields $\phi_{i}$ \cite{cardy2004boundary,cardy1984conformal} and consider $w_{i},\bar w_{i}$ as independent variables. To simplify notation we will write $\mathcal{F}(\left\{w_{i}\right\}_{i=1}^{2N})$.

Now we need to consider the evolution of SLE trace $\gamma_{t}$ from  $t$ to $t+ dt$. First factor in right-hand-side of equation (\ref{eq:2}) gives us
\begin{equation*}
  -\frac{2h_{i}}{w_{i}^{2}}\left(\frac{\partial w_{i}}{\partial z_{i}}\right)^{h_{i}}.
\end{equation*}
We denote by $\mathcal{G}_{i}$ generator of infinitesimal transformation of primary field $\phi_{i}$:$d\phi_{i}(w_{i}) = \mathcal{G}_{i}\phi_{i}(w_{i})$. Then normalization condition of additional $\dim\gf$-dimensional Brownian motion \eqref{eq:5} to $\mathbb{E}\left[d\theta^{a}\; d\theta^{b}\right]=\mathcal{K}(t^{a},t^{b})dt$, and the generator of field transformation \eqref{eq:2} is
\begin{equation}
  \mathcal{G}_{i}=\left(\frac{2dt}{w_{i}}-\sqrt{\kappa} d\xi_{t}\right) \partial_{w_{i}}+\frac{\sqrt{\tau}}{w_{i}}\left(\sum_{a:\mathcal{K}(t^{a},\tilde{t}^{b})=0}\left(d \theta ^{a} t^{a}_{i}\right)\right).
\label{eq:3}
\end{equation}
 Ito formula  gives us the expression for the differential of $\mathcal{F}$, it should be zero due to martingale condition:
\begin{equation}
d \mathcal{F}_{\mathbb{H}_{t}}= \left(\prod_{i=1}^{2N}\left(\frac{\partial w_{i}}{\partial z_{i}}\right)^{h_{i}}\right)\left(-\sum_{i=1}^{2N}\frac{2h_{i}dt}{w_{i}^{2}}+\left[\sum_{i=1}^{2N}\mathcal{G}_{i}+\frac{1}{2}
      \sum_{i,j}\mathcal{G}_{i}\mathcal{G}_{j}\right]\right)\mathcal{F}_{\mathbb{H}}=0
\label{eq:8}
\end{equation}
Substitute definition \eqref{eq:3} and get 
\begin{equation}
  \left(-2 \mathcal{L}_{-2}+\frac{1}{2}\kappa \mathcal{L}_{-1}^{2}+\frac{\tau}{2}\left( \sum_{a} \mathcal{J}^{a}_{-1} \mathcal{J}^{a}_{-1}-
      \sum_{b}\tilde{\mathcal{J}}^{b}_{-1} \tilde{\mathcal{J}}^{b}_{-1}\right)\right)        \mathcal{F}_{\mathbb{H}}=0,
  \label{eq:4}
\end{equation}
where
\begin{equation*}
  \mathcal{L}_{-n}=\sum_{i}\left(\frac{(n-1)h_{i}}{(w_{i}-z)^{n}}-\frac{\partial_{w_{i}}}{(w_{i}-z)^{n-1}}\right);\quad \mathcal{J}^{a}_{{-n}}=-\sum_{i}\frac{t^{a}_{i}}{(w_{i}-z)^{n}};\; \tilde{\mathcal{J}}^{b}_{{-n}}=-\sum_{i}\frac{\tilde{t}^{b}_{i}}{(w_{i}-z)^{n}}
\end{equation*}
This equation is equivalent to following algebraic condition on boundary state $\varphi(0)\left|0\right>$:
\begin{equation}
  \label{eq:7}
  <0|\varphi(\infty)\phi_{1}(w_{1})\dots\phi_{2N}(w_{2N})\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau \left(\sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{-1}-\sum_{b=1}^{\dim\af}\tilde{J}^{b}_{-1}\tilde{J}^{b}_{-1}\right)\right)\varphi(0)|0>=0
\end{equation}
Since the set $\{\phi_{i}\}$ consists of arbitrary primary fields we conclude that $|\psi>=\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau \left(\sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{-1}-\sum_{b=1}^{\dim\af}\tilde{J}^{b}_{-1}\tilde{J}^{b}_{-1}\right)\right)\varphi(0)|0>$ is null-state. Now we act on $\psi$ by raising operators to get equations on $\kappa,\tau$. Since in coset theory
\begin{equation}
  \label{eq:18}
\begin{array}{l}
  L_{n}=L_{n}^{\gf}-L_{n}^{\af}\\
  \left[L_{n}^{\gf},J^{a}_{m}\right]=-m J^{a}_{n+m}\\
  \left[L_{n}^{\gf},\tilde{J}^{b}_{m}\right]=-m\tilde{J}^{b}_{n+m}\\
  \left[L_{n}^{\gf},L_{m}\right]=[L_{n}^{\gf},L_{m}^{\gf}]-[L_{n}^{\gf},L_{m}^{\af}]=(n-m)L_{m+n}+\frac{c}{12}(n^{3}-n)\delta_{m+n,0}
\end{array}
\end{equation}
 it is more convenient to act by $L_{2}^{\gf}$ and $\left(L_{1}^{\gf}\right)^{2}$. 
 Acting with $L_{2}^{\gf}$ we get
\begin{equation*}
  L_{2}^{\gf}\psi= \left(-8 L_{0}-c+ 3 \kappa L_{0}+\frac{1}{2}\tau (k \dim\gf-x_{e}k\dim\af)\right) \varphi_{(\mu,\nu)}=0
\end{equation*}
We use $L_{0} \varphi_{(\mu,\nu)}=h_{(\mu,\nu)} \varphi_{(\mu,\nu)}$ where conformal weight $h_{(\mu,\nu)}= \left(\frac{(\mu,\mu+2\rho)}{2(k+h^{\vee}_{\gf})}-\frac{(\nu,\nu+2\rho_{\af})}{2(k x_{e}+h^{\vee}_{\af})}\right)$ and central charge $c=\frac{k\dim \gf}{k+h^{\vee}_{\gf}}-\frac{x_{e}k\dim \af}{x_{e} k+h^{\vee}_{\af}}$ and get relation on $\kappa,\tau$:
\begin{equation}
  \label{eq:28} (3\kappa-8)h_{(\mu,\nu)}-c+\tau (k\dim\gf-x_{e}k\dim\af) =0
\end{equation}
Second relation appears as the result of $L_{1}^{\gf}$-action:
\begin{equation}
  \label{eq:21}
 -12 h_{(\mu,\nu)}+2\kappa h_{(\mu,\nu)} (2h_{(\mu,\nu)}+1) + \tau
(C_{\mu}-\tilde{C}_{\nu})=0,
\end{equation}
 where $C_{\mu}=(\mu,\mu+2\rho), \tilde{C}_{\nu}=(\nu,\nu+2\rho_{\af})$ is the eigenvalue of quadratic Casimir operators $\sum_{a}t^{a}t^{a}, \sum_{b}\tilde{t}^{b}\tilde{t}^{b}$ of Lie algebras $\gf$ and $\af$.

From equations \eqref{eq:28},\eqref{eq:17} we immediately get $\kappa,\tau$ for each pair $(\mu,\nu)$ of $\gf$ and $\af$-weights.

It is also possible to consider coset theory as $G\times A$  gauge theory and impose constraints on physical states later (See \cite{hwang1993brst,kogan1997knizhnik,gawdzki1988g,figueroa89equivalence}). In this case second term in equation \eqref{eq:3} will include sum over $\gf$ and $\af$ generators and two parameters of Brownian motion $\tau_{\gf}$ and $\tau_{\af}$.

Coset  models are specified by Lie algebra
$\gf$ and its subalgebra $\af$. Denote by $J_{n}^{a}$ the generators of affine Lie algebra $\gfh$
and by $\tilde{J}_{n}^{b}$ the generators of $\afh$, so that $\tilde{J}^{b}_{n}=\sum_{a} m_{a}^{b}
J^{a}_{n}$. 
Virasoro generators in coset models are given by the difference of Sugawara expressions
of $\gf$ and $\af$-WZNW models:
\begin{equation*} L_{n}=L_{n}^{\gf}-L_{n}^{\af}
\end{equation*}
 Commutation relations of Virasoro generators with generators of subalgebra $\afh$
are trivial
\begin{equation}
  \label{eq:26} \left[L_{n},\tilde{J}^{b}_{m}\right]=0
\end{equation}

Commutation relations between generators of $\gfh$ and Virasoro generators are now
\begin{equation}
  \label{eq:1}
  \left[ L_{n}, J_{m}^{a} \right] = [L_{n}^{\gf}-L_{n}^{\af},J_{m}^{a}]= 
\end{equation}

Coset models can be realized as gauged Wess-Zumino-Novikov-Witten models by adding gauge fields
 $A,
\bar{A}$ taking values in Lie algebra $\af\subset \gf$ to the action\cite{gawdzki1988g}.
 Then fields
are labeled by pairs of weights $(\mu,\nu)$, where $\mu$ is the weight of $\gfh$ and $\nu$ is the
weight of $\afh$ correspondingly. But there are selection rules, i.e. the branching functions for
certain pairs of weights $(\mu,\nu)$ vanish. There is also a redundancy, i.e. non-vanishing
branching functions for distinct pairs $(\mu,\nu)$ turn out to be identical
\cite{fuchs1996resolution,schellekens1990field}.

So, primary fields are labeled by pairs of weights $(\mu,\nu)\in \hf_{\gfh}\oplus \hf_{\afh}$ of
algebra $\gfh$ and subalgebra $\afh$, such that branching functions $b^{\mu}_{\nu}(q)\neq 0$. Some
pairs are equivalent. This equivalence is given by the action of so called ``simple currents''
$(J,\tilde{J})$, which are certain elements of outer automorphisms group $\mathcal{O}(\gfh)\times
\mathcal{O}(\afh)\approx B(G)\times B(A)$ of $\gfh\times\afh$, where $B(G)$ is center of Lie group
G. We can think of simple currents as of primary fields and their action on fields of theory are
then given by fusion product \cite{difrancesco1997cft}. So primary fields of coset model are given
by the equivalence classes of pairs of weights $(\mu,\nu)\sim (J*\mu,\tilde{J}*\nu)$, where
$(J,\tilde J)$ such that their conformal weights are equal: $h_{J}-h_{\tilde{J}}=0$.

Conformal weight of coset primary field is equal to
\begin{multline}
L_0\left|\phi_{(\mu,\nu)}\right>=\left(\frac{1}{2(k+h^{\vee})}\sum_aJ^a_0J^a_0-\frac{1}{2(k+h_{\af}^{\vee})}\sum_b
\tilde{J}^b_0 \tilde{J}^b_0 \right) \left|\phi_{(\mu,\nu)}\right>=\\
\left(\frac{(\mu,\mu+2\rho)}{2(k+h^{\vee})}-\frac{(\nu,\nu+2\rho_{\af})}{2(k+h^{\vee})}\right)\left|\phi_{(\mu,\nu)}\right>
\end{multline} It is possible to obtain analogues of Knizhnik-Zamolodchikov equations for coset
models \cite{kogan1997knizhnik}:
\begin{equation} \left\{\frac{1}{2}\partial_{i} + \sum_{i\neq
j}^{N}\left(\frac{t^{a}_{i}t^{a}_{j}}{k+h^{\vee}}-\frac{\tilde t^{b}_{i}\tilde
t^{b}_{j}}{k+h^{\vee}_{\af}}\right)\frac{1}{z_{i}-z_{j}}\right\} \left<\phi_{1}(z_{1})\dots
\phi_{N}(z_{N})\right>=0
  \label{eq:6}
\end{equation}

Let us introduce Schramm-Loewner evolution corresponding to coset models. The idea is to constrain
additional Brownian motion on group manifold to the factor space $G/A$.

To study SLE on $G/A$ factor space we restrict random walk on group manifold by the choice of basis
in $\gf$. Assume that generators $\{J^{a};\; a=1\dots \dim\gf\}$ are chosen in such way that
$\mathcal{K}(J^{a},J^{b})=h^{\vee}\delta_{ab}$ and $\{J^{a};\;
a=\dim\gf-\dim\af\dots \dim\gf\}$ are the generators of subalgebra
$\af\subset \gf$. Now we can consider $(\dim\gf-\dim\af)$-dimensional Brownian
motion with generators $d\theta^{a}$ such that $\mathbb{E}(d\theta^{a} \;
d\theta^{b})=\delta_{ab};\; a,b=1,\dots,\dim\gf-\dim\af$.

Then for generator of field transformation we can write
\begin{equation} \mathcal{G}_{i}=\left(\frac{2dt}{w_{i}}-\sqrt{\kappa}
d\xi_{t}\right) \partial_{w_{i}}+\frac{\sqrt{\tau}}{w_{i}}\left(\sum_{a=1}^{\dim\gf-\dim\af}\left(d
\theta ^{a} t^{a}_{i}\right)\right)
\label{eq:5}
\end{equation}
 Substituting to equation \eqref{eq:8} we get martingale condition

\begin{equation} \left(-2 \mathcal{L}_{-2}+\frac{1}{2}\kappa
\mathcal{L}_{-1}^{2}+\frac{\tau}{2}\left( \sum_{a=1}^{\dim\gf-\dim\af}
\mathcal{J}^{a}_{-1} \mathcal{J}^{a}_{-1}\right)\right) \mathcal{F}_{\mathbb{H}}=0
\label{eq:9}
\end{equation}

which can be rewritten as the requirement for
\begin{equation} \psi=\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau
\left(\sum_{a=1}^{\dim\gf-\dim\af}J^{a}_{-1}J^{a}_{-1}\right)\right)
\phi_{(\mu,\nu)}
\label{eq:10}
\end{equation}
 to be level two null-field.

 It is not always convenient to work with generators
orthogonal with respect to Killing form $\mathcal{K}$, sometimes it is preferable to use Chevalley
or Cartan-Weyl basis. So we change normalization condition of additional
$\dim\gf$-dimensional Brownian motion \eqref{eq:5} to $\mathbb{E}\left[d\theta^{a}\;
d\theta^{b}\right]=\mathcal{K}(t^{a},t^{b})dt$, and the generator of field transformation
\eqref{eq:2} is
\begin{equation*} \mathcal{G}_{i}=\left(\frac{2dt}{w_{i}}-\sqrt{\kappa}
d\xi_{t}\right) \partial_{w_{i}}+\frac{\sqrt{\tau}}{w_{i}}\left(\sum_{a:\mathcal{K}(t^{a},\tilde{t}^{b})=0}\left(d
\theta ^{a} t^{a}_{i}\right)\right).
\end{equation*}
 Martingale condition \eqref{eq:8} can be written as
\begin{equation*} \left(-2 \mathcal{L}_{-2}+\frac{1}{2}\kappa
\mathcal{L}_{-1}^{2}+\frac{\tau}{2}\left( \sum_{a} \mathcal{J}^{a}_{-1} \mathcal{J}^{a}_{-1}-
\sum_{b}\tilde{\mathcal{J}}^{b}_{-1} \tilde{\mathcal{J}}^{b}_{-1}\right)\right)
\mathcal{F}_{\mathbb{H}}=0.
\end{equation*}
 This equation can again be rewritten as the requirement for
\begin{equation} \psi=\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau
\left(\sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{-1}-\sum_{b=1}^{\dim\af}\tilde{J}^{b}_{-1}\tilde{J}^{b}_{-1}\right)\right)
\phi_{(\mu,\nu)}
\label{eq:30}
\end{equation}
 to be level two null-field.

Acting with $L_{2}$ we get
\begin{equation*}
  L_{2}\psi= \left(-8 L_{0}+c/2+ 3 \kappa L_{0}+\frac{1}{2}\tau k \dim\gf\right)
\phi_{(\mu,\nu)}
\end{equation*}
\begin{equation}
  \label{eq:28} (3\kappa-8) h_{(\mu,\nu)}+c+\tau k =0
\end{equation}

Now we act with $L_{1}^{2}$  and obtain
\begin{equation}
  \label{eq:17} 12 h_{(\mu,\nu)}+2\kappa h_{(\mu,\nu)} (2h_{(\mu,\nu)}+1) + \tau
(C_{\mu}-\tilde{C}_{\nu})=0,
\end{equation}
 where $C_{\mu}=(\mu,\mu+2\rho)$ is the eigenvalue of quadratic Casimir operator
$\sum_{a}t^{a}t^{a}$. We need to get one more equation since we
have three variables $\kappa,\tau,h_{(\mu,\nu)}$. We can use that $\psi$ is singular vector so
$L_{1}\psi=0$ and then $J_{1}^{a}L_{1}\psi=0,\; \tilde{J}_{1}^{b} L_{1}\psi=0$, so

\begin{equation}
L_{1} \psi=L_{1}\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau
\left(\sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{-1}-\sum_{b=1}^{\dim\af}\tilde{J}^{b}_{-1}\tilde{J}^{b}_{-1}\right)\right)
\phi_{(\mu,\nu)}
\end{equation}

\begin{equation*}
L_{1}\psi = \left(-6 L_{-1}+\kappa (2  L_{-1} L_{0}+L_{0}) + \tau \sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{0}\right)
\phi_{(\mu,\nu)}
\end{equation*}
\begin{equation*}
  L_{1}^{2}\psi= \left(-12 L_{0}+\kappa (4  L_{0}^{2}+2 L_{0}) + \tau \sum_{a=1}^{\dim\gf}J^{a}_{0}J^{a}_{0}\right)
\phi_{(\mu,\nu)}
\end{equation*}
\begin{equation*}
  J_{1}^{a}L_{1} \psi= \left(-6 J_{0}^{a}+\kappa (2  J_{0}^{a} L_{0}+J_{0}^{a}) + \tau
    \sum_{b=1}^{\dim\gf}\sum_{c=1}^{\dim\gf}i f^{abc}J^{c}_{0}J^{b}_{0}+k\tau J_{0}^{a}\right)
\phi_{(\mu,\nu)}
\end{equation*}

\begin{equation*}
  L_{2}\psi= \left(-8 L_{0}+c/2+ 3 \kappa L_{0}+\frac{1}{2}\tau k \dim\gf\right)
\phi_{(\mu,\nu)}
\end{equation*}

\begin{eqnarray*}
  (3\kappa-8) h_{(\mu,\nu)}+c/2+\frac{1}{2}\tau k \dim\gf=0\\
  -12 h_{(\mu,\nu)}+\kappa (4 h_{(\mu,\nu)}^{2}+2 h_{(\mu,\nu)}) +\tau C(\mu) =0\\
  (-6 +\kappa (2 h_{(\mu,\nu)}+1)+\tau
\end{eqnarray*}
\begin{equation}
  \label{eq:29} \left(-6 + \kappa + \tau + 2 \kappa h_{(\mu,\nu)}\right) J^{3}_{0}
\phi_{(\mu,\nu)}=0
\end{equation}

Acting on \eqref{eq:30} with $L_{2}$ and $L_{1}^{2}$ we can obtain equations \eqref{eq:28} and
\eqref{eq:17}. To get more equations we can act by $\afh$-generators $\tilde{J}^{b}_{1}$ and $L_{1}$
similarly to what was done with $J^{3}_{1}$ to get equation \eqref{eq:29}.

To obtain simpler relations it is also possible to use generalization of Knizhnik-Zamolodchikov
equations \eqref{eq:6} as was done for WZNW-models in paper \cite{alekseev2010sle}.

Now consider simple example. Let $G=SU(2)$ and $A=U(1)$, and corresponding Lie algebras
$\gf=\mathrm{su}(2)$ with generators $J^{1},J^{2},J^{3}$ and $\af=\mathrm{u}(1)$ with the generator
$J^{3}$, $\af\subset\gf$. Note that $\mathcal{K}(J^{a},J^{b})=2\delta^{ab}$.

 The equation \eqref{eq:10} is now
\begin{equation}
  \label{eq:11} \psi=\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau
\left(J^{1}_{-1}J^{1}_{-1}+J^{2}_{-1}J^{2}_{-1}\right)\right) \phi_{(\mu,\nu)}
\end{equation}
 If we use basis $J^{+}=\frac{J^{1}+iJ^{2}}{\sqrt{2}},\;
J^{-}=\frac{J^{1}-iJ^{2}}{\sqrt{2}}$ this equation is rewritten in the form
\begin{equation} \psi= \left(-2
L_{-2}+\frac{\kappa}{2}L_{-1}^{2}+\frac{\tau}{2}\left[J^{+}_{-1}J^{-}_{-1}+J^{-}_{-1}J^{+}_{-1}\right]\right)
\phi_{(\mu,\nu)}
\label{eq:12}
\end{equation}
 which is similar to the equations for parafermionic fields introduced in paper
\cite{santachiara2008sle}.

Central charge in this case is equal to
\begin{equation}
  \label{eq:14} c=\frac{2(k-1)}{k+2}
\end{equation}

Let us study the solutions of the equation \eqref{eq:11}. Act on $\psi$ with $L_{2}$:
\begin{equation}
  \label{eq:13} L_{2}\psi=(-8L_{0}+c+3\kappa L_{0}+\tau k)\phi_{(\mu,\nu)}=0,
\end{equation}
 and obtain
\begin{equation}
  \label{eq:28} (3\kappa-8) h_{(\mu,\nu)}+c+\tau k =0
\end{equation}
 Another equation can be obtained by action of $L_{1}^{2}$
\begin{equation}
  \label{eq:15} L_{1}^{2}\psi = (12 L_{0} + \kappa(4 L_{0}^{2}+2 L_{0}) +\tau
(J_{0}^{1}J_{0}^{1}+J_{0}^{2}J_{0}^{2}))\phi_{(\mu,\nu)}=0
\end{equation}
 Since
$J_{0}^{1}J_{0}^{1}+J_{0}^{2}J_{0}^{2}=J_{0}^{1}J_{0}^{1}+J_{0}^{2}J_{0}^{2}+J_{0}^{3}J_{0}^{3}-J_{0}^{3}J_{0}^{3}$
we have
\begin{equation}
  \label{eq:17} 12 h_{(\mu,\nu)}+2\kappa h_{(\mu,\nu)} (2h_{(\mu,\nu)}+1) + \tau
(C_{\mu}-\tilde{C}_{\nu})=0,
\end{equation}
 where $C_{\mu}=(\mu,\mu+2\rho)$ is the eigenvalue of quadratic Casimir operator
$\sum_{a}t^{a}t^{a}$. For $u(1)$ it is just $\nu^{2}$. We need to get one more equation since we
have three variables $\kappa,\tau,h_{(\mu,\nu)}$. We can use that $\psi$ is singular vector so
$L_{1}\psi=0$ and then $J_{1}^{3}L_{1}\psi=0$, so
\begin{equation}
  \label{eq:29} \left(-6 + \kappa + \tau + 2 \kappa h_{(\mu,\nu)}\right) J^{3}_{0}
\phi_{(\mu,\nu)}=0
\end{equation}

Turn to the case $k=2, \;c=1/2$ which corresponds to square lattice Ising model. We have three
non-equivalent fields numbered by the pairs of $su(2)$ and $u(1)$ weights, which are $\phi_{(0,0)},
\; h_{(0,0)}=0; \quad \phi_{(0,2)}, \; h_{(0,2)}=1/2; \quad \phi_{(1,1)}, \; h_{(1,1)}=1/16$. The
equation \eqref{eq:29} is trivial for $\phi_{(0,0)}$ and $\phi_{(0,2)}$, so the equations
\eqref{eq:17}, \eqref{eq:28} are consistent although we do not know how to interpret $\tau$ in
context of the Ising model. If we consider the field $\phi_{(1,1)}$ we have no solution for
$\kappa,\tau$, so we see that not every primary field corresponds to some boundary condition
changing operator.

\section{Algebraic equations}
\label{sec:algebraic-equations}

\section{Examples}
\label{sec:examples}

\section{Conclusion}
\label{sec:conclusion}

\bibliography{bibliography}{}
\bibliographystyle{utphys}

\end{document}