### 2012-04-25 AffineLieAlgebras/SLE/paper.tex
Some changes in SLE paper. Added Smirnov-Makarov paper to bibliography.

\documentclass{jetpl}
\twocolumn

%%% article in English
\lat

%%% declaration of a new mathematical operator
\DeclareMathOperator{\sign}{sign}

%\usepackage{amsmath,amssymb,amsthm}
%\usepackage{multicol}
%\usepackage{color}
%\usepackage{graphicx}
\usepackage{hyperref}

\newtheorem{statement}{Statement}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}{Lemma}
\newtheorem{mynote}{Note}[section]
%\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newcommand{\go}{\stackrel{\circ }{\mathfrak{g}}}
\newcommand{\ao}{\stackrel{\circ }{\mathfrak{a}}}
\newcommand{\co}[1]{\stackrel{\circ }{#1}}
\newcommand{\pia}{\pi_{\mathfrak{a}}}
\newcommand{\piab}{\pi_{\mathfrak{a}_{\bot}}}
\newcommand{\gf}{\mathfrak{g}}
\newcommand{\af}{\mathfrak{a}}
\newcommand{\aft}{\widetilde{\mathfrak{a}}}
\newcommand{\afb}{\mathfrak{a}_{\bot}}
\newcommand{\hf}{\mathfrak{h}}
\newcommand{\hfb}{\mathfrak{h}_{\bot}}
\newcommand{\pf}{\mathfrak{p}}
\newcommand{\gfh}{\hat{\mathfrak{g}}}
\newcommand{\afh}{\hat{\mathfrak{a}}}
\newcommand{\bff}{\mathfrak{b}}
\newcommand{\hfg}{\hf_{\gf}}

%\theoremstyle{definition} \newtheorem{Def}{Definition}
\newcommand{\adk}{a^{\dagger}_{\kappa}} \newcommand{\ak}{a_{\kappa}}
\def\bF{\mbox{$\overline{\cal F}$}} \def\F{\mbox{$\cal F$}}

%%% article title
\title{SLE martingales in coset conformal field theory}

%%% article title - for colontitle (at the top of the page)
\rtitle{SLE martingales in coset conformal field theory}

%%% article title - for table of contents (usualy identical with \title)
\sodtitle{SLE martingales in coset conformal field theory}

%%% author(s) ( + e-mail)

\author{Anton Nazarov$^{+*}$\/\thanks{e-mail: anton.nazarov@hep.phys.spbu.ru}}

%%% author(s) - for colontitle (at the top of the page)
\rauthor{Anton Nazarov}

%%% author(s) - for table of contents
\sodauthor{Nazarov}

%%% author's address(es)
\address{
  $^+$ Department of High-Energy and Elementary Particle Physics, 
  Faculty of Physics and\\ Chebyshev Laboratory,
  Faculty of Mathematics and Mechanics, \\ SPb State University
  198904, Saint-Petersburg, Russia}

%%% dates of submition & resubmition (if submitted once, second argument is *)
\dates{06 April 2012}{*}

%%% abstract

\abstract{
  Scharmm-Loewner evolution (SLE) and conformal field theory (CFT) are popular instruments to study critical behavior of physical models in two dimensions, but they use different objects. While SLE  has natural connection with lattice models and is suitable for strict proofs, it lacks computational and predictive power of conformal field theory.
  To provide a way for the concurrent use of SLE and CFT we consider CFT correlation functions which are martingales with respect to SLE.  We establish necessary and sufficient conditions on such observables for coset models of conformal field theory and Schramm-Loewner evolution on group manifolds and check the consistency with the  behaviour of parafermionic models. Coset models are connected with off-critical massive field theories and we discuss possible implications for SLE.}

%%% PACS numbers
\PACS{05.50.+q, 11.25.Hf, 02.20.Sv}

\begin{document}
\maketitle

\section{Introduction}
\label{sec:introduction}

Schramm-Loewner evolution \cite{schramm2000scaling} is a stochastic process satisfying stochastic differential equation (\ref{eq:19}). Solution of equation (\ref{eq:19}) can be seen as probability measure on random curves called SLE traces. SLE is  useful for the study of critical behavior \cite{Cardy:2005kh,rohde2005basic}, since for many lattice models  the convergence of  domain walls to SLE traces can be proved \cite{bauer20062d,schramm2006conformally}. Practical computations with SLE rely on Ito calculus.

Conformal field theory \cite{belavin1984ics} is formulated in terms of families of quantum fields which usually has no natural interpretation in microscopic terms. On the other hand CFT has efficient computational tools such as Virasoro symmetry, Knizhnik-Zamolodchikov equations, fusion algebra etc and provides numerous predictions, such as celebrated Cardy formula \cite{cardy1992critical,smirnov2001critical} which gives crossing probability for critical percolation.

Combination of these two approaches should be immensely powerful, so it is important to find objects which can be studied from both points of view. We consider correlation functions of CFT which are martingales with respect to Schramm-Loewner evolution are such objects.

We present new class of such observables in coset conformal field theory \cite{Goddard198588}. They are correlation functions in $G/A$-coset conformal field theory which are martingales with respect to Schramm-Loewner evolution with additional Brownian motion on factor-space of Lie group $G$ by subgroup $A$ \cite{2011arXiv1112.4354N}. Using Knizhnik-Zamolodchikov equations we establish necessary and sufficient conditions on such functions and derive relations connecting parameters of SLE and CFT. We then check the consistency of these results with the earlier results obtained for parafermionic observables \cite{santachiara2008sle}.

All minimal unitary models can be obtained by coset construction. This coset structure leads to specific off-critical behavior such as classification of massive excitations by affine Toda field theory \cite{fateev1990conformal,eguchi1989deformations,hollowood1989rational}. In conclusion we discuss possible connections of these CFT perturbation with massive off-critical SLEs (which contain additional drift term  \cite{makarov2010off}).

\section{Martingale conditions}
\label{sec:mart-cond}

%% We
%% consider Schramm-Loewner evolution $SLE_{\kappa}$ with additional Brownian motion on coset space
%% $G/A$ characterised by parameter $\tau$ proposed in our paper \cite{2011arXiv1112.4354N}. We consider correlation functions in coset models of
%% conformal field theory which satisfy martingale conditions with respect to this stochastic process.
%% Since all minimal unitary models of conformal field theory can be obtained from coset construction
%% for certain choice of Lie group $G$ and subgroup $A$ \cite{difrancesco1997cft} these correlation functions correspond to SLE
%% observables appearing as the scaling limit of lattice observables. Trace of
%% Schramm-Loewner $\gamma_{t}$ evolution describes critical domain wall in lattice models.
%% 
 Consider some critical lattice model on the upper half-plane $\mathbb{H}$ with the cut along a critical interface $\gamma_{t}$. We denote this slit domain by $\mathbb{H}_{t}=\mathbb{H}\setminus \gamma_{t}$.  Assume that $\gamma_{t}$ satisfies restriction property and is conformally  invariant \cite{Cardy:2005kh}. Then  conformal map  $g_{t}:\mathbb{H}_{t}\to \mathbb{H}$ satisfies stochastic differential equation \cite{schramm2000scaling}:
\begin{equation}
\label{eq:19}
  \frac{\partial g_t(z)}{\partial t} = \frac{ 2}{g_t(z)-\sqrt{\kappa}\xi_{t}} ,
\end{equation}
where $\xi_{t}$ is the Brownian motion. The dynamic of the tip $z_{t}$ of critical curve $\gamma_{t}$ (tip of SLE trace) is given by the law $z_{t}=g_{t}^{-1}(\sqrt{\kappa}\xi_{t})$.

For us it is more convenient to use map $w_{t} (z)=g_{t}(z)-\sqrt{\kappa}\xi_{t}$, so the equation \eqref{eq:19} becomes
\begin{equation}
  \label{eq:20}
       d w _{t}= \frac{2dt}{w_{t} }-\sqrt{\kappa}\xi_{t}  
\end{equation}

Consider  $N$-point correlation function of boundary conformal field theory with boundary condition changing operators sitting at the tip of SLE trace and at the infinity:
\begin{equation}
  \mathcal{F}(\left\{z_{i},\bar z_{i}\right\}_{i=1}^{N})_{\mathbb{H}_{t}}=
\left<\varphi(z_{t}) \phi_{1}(z_{1},\bar z_{1}) \dots \phi_{n}(z_{n},\bar z_{n})
      \varphi(\infty)\right>,
    \label{eq:2}
\end{equation}
where $\phi_{i}$ are bulk primary fields and $\varphi(z_{t})$ is boundary condition changing operator.

Primary fields in coset models are labeled by pairs $(\lambda,\eta)$ of weights $\gf$ and $\af$-representations. For each $\lambda$ possible values of $\eta$ are those which appear in decomposition of representation $L^{(\mu)}_{\gf}=\bigoplus_{\eta} H^{\lambda}_{\eta}\otimes L^{(\eta)}_{\af}$. Some pairs are equivalent, we discuss and use this equivalence relation later. 
We assume that boundary field $\varphi$ is also primary and labeled by weights $(\mu,\nu)$.

  We use the conformal map  $w(z):\mathbb{H}\setminus\gamma_{t}\to \mathbb{H}$ to rewrite expression \eqref{eq:2} in the whole upper half plane:
\begin{multline}
  \mathcal{F}(\left\{z_{i},\bar z_{i}\right\})_{\mathbb{H}_{t}}=\\
  =\prod \left(\frac{\partial w(z_{i})}{\partial z_{i}}\right)^{h_{i}} 
  \prod \left(\frac{\partial \bar w(\bar z_{i})}{\partial \bar z_{i}}\right)^{\bar h_i}
  \mathcal{F}(\left\{w_{i}, \bar w_{i}\right\})_{\mathbb{H}}
  \label{eq:1}
\end{multline}
We can extend the definition of $\mathcal{F}$ to the whole plane by doubling the number of fields $\phi_{i}$ \cite{cardy2004boundary,cardy1984conformal} and consider $w_{i},\bar w_{i}$ as independent variables. To simplify notation we will write $\mathcal{F}(\left\{w_{i}\right\}_{i=1}^{2N})$.

$G/A$-coset conformal field theory can be realized as WZNW-model with gauge group $G$ interacting with pure gauge field of gauge group $A\subset G$ \cite{gawdzki1988g,figueroa89equivalence}. So we have $G/A$ gauge symmetry and we add random gauge transformations to Schramm-Loewner evolution \cite{2011arXiv1112.4354N} similar to the case of WZNW-models \cite{bettelheim2005stochastic}. Denote by $\mathcal{K}$ Killing form of Lie algebra $\gf$ corresponding to Lie group $G$ and by $t^{a}_{i}$ ($\tilde{t}^{b}_{i}$) the  generators of $\gf$-representation ($\af$-representation) corresponding to primary field $\phi_{i}$.

Now we need to consider the evolution of SLE trace $\gamma_{t}$ from  $t$ to $t+ dt$. First factor in right-hand-side of equation (\ref{eq:2}) gives us
\begin{equation*}
  -\frac{2h_{i}}{w_{i}^{2}}\left(\frac{\partial w_{i}}{\partial z_{i}}\right)^{h_{i}}.
\end{equation*}
We denote by $\mathcal{G}_{i}$ generator of infinitesimal transformation of primary field $\phi_{i}$:$d\phi_{i}(w_{i}) = \mathcal{G}_{i}\phi_{i}(w_{i})$. We normalize additional $\dim\gf$-dimensional Brownian motion as $\mathbb{E}\left[d\theta^{a}\; d\theta^{b}\right]=\mathcal{K}(t^{a},t^{b})dt$, and the generator of field transformation is equal to 
\begin{equation}
  \mathcal{G}_{i}=\left(\frac{2dt}{w_{i}}-\sqrt{\kappa} d\xi_{t}\right) \partial_{w_{i}}+\frac{\sqrt{\tau}}{w_{i}}\left(\sum_{a:\mathcal{K}(t^{a},\tilde{t}^{b})=0}\left(d \theta ^{a} t^{a}_{i}\right)\right).
\label{eq:3}
\end{equation}
 Ito formula  gives us the expression for the differential of $\mathcal{F}$ which should be zero due to martingale condition:
\begin{multline}
d \mathcal{F}_{\mathbb{H}_{t}}=\\ \left(\prod_{i=1}^{2N}\left(\frac{\partial w_{i}}{\partial z_{i}}\right)^{h_{i}}\right)
\left(-\sum_{i=1}^{2N}\frac{2h_{i}dt}{w_{i}^{2}}+\left[\sum_{i=1}^{2N}\mathcal{G}_{i}+\frac{1}{2}
      \sum_{i,j}\mathcal{G}_{i}\mathcal{G}_{j}\right]\right)\mathcal{F}_{\mathbb{H}}\\=0
\label{eq:8}
\end{multline}
Substitute definition \eqref{eq:3} and get 
\begin{equation}
  \left(-2 \mathcal{L}_{-2}+\frac{1}{2}\kappa \mathcal{L}_{-1}^{2}+\frac{\tau}{2}\left( \sum_{a} \mathcal{J}^{a}_{-1} \mathcal{J}^{a}_{-1}-
      \sum_{b}\tilde{\mathcal{J}}^{b}_{-1} \tilde{\mathcal{J}}^{b}_{-1}\right)\right)        \mathcal{F}_{\mathbb{H}}=0,
  \label{eq:4}
\end{equation}
where
\begin{eqnarray*}
  \mathcal{L}_{-n}=\sum_{i}\left(\frac{(n-1)h_{i}}{(w_{i}-z)^{n}}-\frac{\partial_{w_{i}}}{(w_{i}-z)^{n-1}}\right)\\ \mathcal{J}^{a}_{{-n}}=-\sum_{i}\frac{t^{a}_{i}}{(w_{i}-z)^{n}};\; \tilde{\mathcal{J}}^{b}_{{-n}}=-\sum_{i}\frac{\tilde{t}^{b}_{i}}{(w_{i}-z)^{n}}
\end{eqnarray*}
This equation is equivalent to following algebraic condition on boundary state $\varphi(0)\left|0\right>$:
\begin{multline}
  \label{eq:7}
  \left<0\left|\varphi(\infty)\phi_{1}(w_{1})\dots\phi_{2N}(w_{2N})\right.\right.\\
  \left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau \left(\sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{-1}-\sum_{b=1}^{\dim\af}\tilde{J}^{b}_{-1}\tilde{J}^{b}_{-1}\right)\right)\\
\left.\varphi(0)|0\right>=0
\end{multline}
Since the set $\{\phi_{i}\}$ consists of arbitrary primary fields we conclude that 
\begin{multline}
|\psi>=\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau \left(\sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{-1}-\sum_{b=1}^{\dim\af}\tilde{J}^{b}_{-1}\tilde{J}^{b}_{-1}\right)\right)\\
\cdot\varphi(0)|0>
\end{multline}
is null-state. Now we act on $\psi$ by raising operators to get equations on $\kappa,\tau$. Since in coset theory
\begin{equation}
  \label{eq:18}
\begin{array}{ll}
  L_{n}= L_{n}^{\gf}-L_{n}^{\af} &\\
  \left[L_{n}^{\gf},J^{a}_{m}\right]= -m J^{a}_{n+m}&\\
  \left[L_{n}^{\gf},\tilde{J}^{b}_{m}\right]=-m\tilde{J}^{b}_{n+m}&\\
  \left[L_{n}^{\gf},L_{m}\right]=[L_{n}^{\gf},L_{m}^{\gf}]-[L_{n}^{\gf},L_{m}^{\af}]=&\\
  \quad =(n-m)L_{m+n}+\frac{c}{12}(n^{3}-n)\delta_{m+n,0}&
\end{array}
\end{equation}
 it is more convenient to act by $L_{2}^{\gf}$ and $\left(L_{1}^{\gf}\right)^{2}$. 
 Acting with $L_{2}^{\gf}$ we get
\begin{equation*}
  L_{2}^{\gf}\psi= \left(-8 L_{0}-c+ 3 \kappa L_{0}+\frac{1}{2}\tau (k \dim\gf-x_{e}k\dim\af)\right) \varphi_{(\mu,\nu)}=0
\end{equation*}
We use $L_{0} \varphi_{(\mu,\nu)}=h_{(\mu,\nu)} \varphi_{(\mu,\nu)}$ where conformal weight $h_{(\mu,\nu)}= \left(\frac{(\mu,\mu+2\rho)}{2(k+h^{\vee}_{\gf})}-\frac{(\nu,\nu+2\rho_{\af})}{2(k x_{e}+h^{\vee}_{\af})}\right)$ and central charge $c=\frac{k\dim \gf}{k+h^{\vee}_{\gf}}-\frac{x_{e}k\dim \af}{x_{e} k+h^{\vee}_{\af}}$ and get relation on $\kappa,\tau$:
\begin{equation}
  \label{eq:28} (3\kappa-8)h_{(\mu,\nu)}-c+\tau (k\dim\gf-x_{e}k\dim\af) =0
\end{equation}
Second relation appears as the result of $L_{1}^{\gf}$-action:
\begin{equation}
  \label{eq:21}
 -12 h_{(\mu,\nu)}+2\kappa h_{(\mu,\nu)} (2h_{(\mu,\nu)}+1) + \tau
(C_{\mu}-\tilde{C}_{\nu})=0,
\end{equation}
 where $C_{\mu}=(\mu,\mu+2\rho), \tilde{C}_{\nu}=(\nu,\nu+2\rho_{\af})$ is the eigenvalue of quadratic Casimir operators $\sum_{a}t^{a}t^{a}, \sum_{b}\tilde{t}^{b}\tilde{t}^{b}$ of Lie algebras $\gf$ and $\af$.
Relations \eqref{eq:28},\eqref{eq:21} are necessary conditions for CFT correlation functions to be SLE martingales.

From equations \eqref{eq:28},\eqref{eq:21} we immediately get $\kappa,\tau$ for each pair $(\mu,\nu)$ of $\gf$ and $\af$-weights.

\subsection{Examples}
\label{sec:examples-1}

Minimal models $su(2)/u(1)$, $su(2)\oplus su(2)/su(2)$.

Consider standard coset realization of minimal models: $su(2)_{N}/u(1)_{N}$. Central charge is $c=\frac{3N}{N+2}-1=\frac{2N-2}{N+2}$. Conformal weights of primary fields with Dynkin indeces $(k,l)$ are $h_{(k,l)}=\frac{k(k+2)}{4(N+2)}-\frac{l^{2}}{4N}$. $C_{\mu}=k(k+2)/2, C_{\nu}=l^{2}/2$
For Ising model $N=2,c=1/2$, $h_{(2,0)}=1/2, h_{(1,1)}=1/16$. We get for $\varphi_{(2,0)}$: $3\kappa-9+4\tau =0, -3+\kappa+4\tau=0$, $\kappa=3, \tau=0$. For $\varphi_{(1,1)}$: $3\kappa-16+64\tau=0, -64+9\kappa + 64\tau=0$, $\kappa=16/3, \tau=0$. 
For $N=3, c=4/5$. Conformal weights $h_{(0,0)}=0$, $h_{(0,2)}=2/3$, $h_{(0,-2)}=2/3$, $h_{(1,3)}=2/5$, $h_{(1,1)}=1/15$, $h_{(1,-1)}=1/15$.  Corresponding values of $\kappa,\tau$ are: $(208/25,242/225), (460/33,116/33),(260/53,28/477)$

Now use another coset realization of minimal unitary models: $\frac{\hat{su}(2)_{N}\oplus \hat{su}(2)_{1}}{\hat{su}(2)_{N+1}}$, $c=\frac{3N}{N+2}+1-\frac{3(N+1)}{N+3}=1-\frac{6}{(N+2)(N+3)}$

three-state Potts model $\frac{\hat{su}(3)_{1}\oplus \hat{su}(3)_{1}}{\hat{su}(3)_{2}}$

But correlation functions in coset conformal field theory must satisfy (generalized) Knizhnik-Zamolodchikov equations \cite{kogan1997knizhnik} which are of the form
\begin{equation}
  \label{eq:16}
 \left[ \frac{1}{2}\frac{\partial}{\partial w_{i}} +\sum_{j\neq i} \left( \frac{\sum_{a}t_{i}^{a}t_{j}^{a}}{k+h^{\vee}}-\frac{\sum_{b}\tilde t_{i}^{b}\tilde t_{j}^{b}}{x_{e}k+h_{\af}^{\vee}}\right) \right]\mathcal{F}_{\mathbb{H}}=0
\end{equation}
Rewrite the equation as
\begin{equation}
  \label{eq:22}
  \left[ \frac{1}{2}\frac{\partial}{\partial w_{i}} +\sum_{j\neq i} \left( \frac{\sum_{a}t_{i}^{a}t_{j}^{a}}{k+h^{\vee}}-\frac{\sum_{b}\tilde t_{i}^{b}\tilde t_{j}^{b}}{x_{e}k+h_{\af}^{\vee}}\right) \right]\mathcal{F}_{\mathbb{H}}=0
\end{equation}

From this relations and equation \eqref{eq:8} we can write sufficient conditions for SLE martingale to be CFT correlation function:
\begin{equation}
  \label{eq:23}
  M x = 0
\end{equation}

It is also possible to consider coset theory as $G\times A$  gauge theory and impose constraints on physical states later (See \cite{hwang1993brst,kogan1997knizhnik,gawdzki1988g,figueroa89equivalence}). In this case second term in equation \eqref{eq:3} will include sum over $\gf$ and $\af$ generators and two parameters of Brownian motion $\tau_{\gf}$ and $\tau_{\af}$.

\section{Conclusion and outlook}
\label{sec:outlook}

In excellent paper \cite{bettelheim2005stochastic} connection between WZNW-models and Schramm-Loewner evolution with additional Brownian motion on group manifold was established. This paper also stated a problem of the connection of SLE parameters for martingales in WZNW, coset and minimal models. In present letter we used the method to obtain necessary and sufficient conditions on SLE martingales proposed in paper \cite{alekseev2010sle}. This method allowed us to compare our results with parafermionic results of the papers \cite{santachiara2008sle,picco2008numerical}.

Coset structure of minimal models is manifest in field theory perturbed by external magnetic field \cite{fateev1990conformal,eguchi1989deformations,hollowood1989rational}. This theory is supported by experimental data \cite{coldea2010quantum}. Relation between correlation functions in coset theory and SLE observables can be a starting point for the study of domain walls in lattice models away of critical point.

Massive off-critical SLEs has additional drift term in driving Brownian motion \cite{makarov2010off}.

  Schramm-Loewner evolution was applied for the description of domain walls in Ising model with external field in recent paper \cite{stevenson2011domain}, although the perturbation was not compatible with affine field theory structure.

\section*{Acknowledgements}
\label{sec:acknowledgements}
 I am grateful to A. Bytsko, K. Izyurov, D. Chelkak and R. Santachiara for the discussions on SLE and WZNW models. This work  is supported by
the Chebyshev Laboratory (Department of Mathematics and Mechanics,
Saint-Petersburg State University) under the grant 11.G34.31.0026
of the Government of the Russian Federation.

%%  
%%  
%%  
%%  Coset  models are specified by Lie algebra
%%  $\gf$ and its subalgebra $\af$. Denote by $J_{n}^{a}$ the generators of affine Lie algebra $\gfh$
%%  and by $\tilde{J}_{n}^{b}$ the generators of $\afh$, so that $\tilde{J}^{b}_{n}=\sum_{a} m_{a}^{b}
%%  J^{a}_{n}$. 
%%  Virasoro generators in coset models are given by the difference of Sugawara expressions
%%  of $\gf$ and $\af$-WZNW models:
%%  \begin{equation*} L_{n}=L_{n}^{\gf}-L_{n}^{\af}
%%  \end{equation*}
%%   Commutation relations of Virasoro generators with generators of subalgebra $\afh$
%%  are trivial
%%  \begin{equation}
%%    \label{eq:26} \left[L_{n},\tilde{J}^{b}_{m}\right]=0
%%  \end{equation}
%%  
%%  Commutation relations between generators of $\gfh$ and Virasoro generators are now
%%  \begin{equation}
%%    \label{eq:1}
%%    \left[ L_{n}, J_{m}^{a} \right] = [L_{n}^{\gf}-L_{n}^{\af},J_{m}^{a}]= 
%%  \end{equation}
%%  
%%  Coset models can be realized as gauged Wess-Zumino-Novikov-Witten models by adding gauge fields
%%   $A,
%%  \bar{A}$ taking values in Lie algebra $\af\subset \gf$ to the action\cite{gawdzki1988g}.
%%   Then fields
%%  are labeled by pairs of weights $(\mu,\nu)$, where $\mu$ is the weight of $\gfh$ and $\nu$ is the
%%  weight of $\afh$ correspondingly. But there are selection rules, i.e. the branching functions for
%%  certain pairs of weights $(\mu,\nu)$ vanish. There is also a redundancy, i.e. non-vanishing
%%  branching functions for distinct pairs $(\mu,\nu)$ turn out to be identical
%%  \cite{fuchs1996resolution,schellekens1990field}.
%%  
%%  So, primary fields are labeled by pairs of weights $(\mu,\nu)\in \hf_{\gfh}\oplus \hf_{\afh}$ of
%%  algebra $\gfh$ and subalgebra $\afh$, such that branching functions $b^{\mu}_{\nu}(q)\neq 0$. Some
%%  pairs are equivalent. This equivalence is given by the action of so called ``simple currents''
%%  $(J,\tilde{J})$, which are certain elements of outer automorphisms group $\mathcal{O}(\gfh)\times
%%  \mathcal{O}(\afh)\approx B(G)\times B(A)$ of $\gfh\times\afh$, where $B(G)$ is center of Lie group
%%  G. We can think of simple currents as of primary fields and their action on fields of theory are
%%  then given by fusion product \cite{difrancesco1997cft}. So primary fields of coset model are given
%%  by the equivalence classes of pairs of weights $(\mu,\nu)\sim (J*\mu,\tilde{J}*\nu)$, where
%%  $(J,\tilde J)$ such that their conformal weights are equal: $h_{J}-h_{\tilde{J}}=0$.
%%  
%%  Conformal weight of coset primary field is equal to
%%  \begin{multline}
%%  L_0\left|\phi_{(\mu,\nu)}\right>=\left(\frac{1}{2(k+h^{\vee})}\sum_aJ^a_0J^a_0-\frac{1}{2(k+h_{\af}^{\vee})}\sum_b
%%  \tilde{J}^b_0 \tilde{J}^b_0 \right) \left|\phi_{(\mu,\nu)}\right>=\\
%%  \left(\frac{(\mu,\mu+2\rho)}{2(k+h^{\vee})}-\frac{(\nu,\nu+2\rho_{\af})}{2(k+h^{\vee})}\right)\left|\phi_{(\mu,\nu)}\right>
%%  \end{multline} It is possible to obtain analogues of Knizhnik-Zamolodchikov equations for coset
%%  models \cite{kogan1997knizhnik}:
%%  \begin{equation} \left\{\frac{1}{2}\partial_{i} + \sum_{i\neq
%%  j}^{N}\left(\frac{t^{a}_{i}t^{a}_{j}}{k+h^{\vee}}-\frac{\tilde t^{b}_{i}\tilde
%%  t^{b}_{j}}{k+h^{\vee}_{\af}}\right)\frac{1}{z_{i}-z_{j}}\right\} \left<\phi_{1}(z_{1})\dots
%%  \phi_{N}(z_{N})\right>=0
%%    \label{eq:6}
%%  \end{equation}
%%  
%%  Let us introduce Schramm-Loewner evolution corresponding to coset models. The idea is to constrain
%%  additional Brownian motion on group manifold to the factor space $G/A$.
%%  
%%  
%%  To study SLE on $G/A$ factor space we restrict random walk on group manifold by the choice of basis
%%  in $\gf$. Assume that generators $\{J^{a};\; a=1\dots \dim\gf\}$ are chosen in such way that
%%  $\mathcal{K}(J^{a},J^{b})=h^{\vee}\delta_{ab}$ and $\{J^{a};\;
%%  a=\dim\gf-\dim\af\dots \dim\gf\}$ are the generators of subalgebra
%%  $\af\subset \gf$. Now we can consider $(\dim\gf-\dim\af)$-dimensional Brownian
%%  motion with generators $d\theta^{a}$ such that $\mathbb{E}(d\theta^{a} \;
%%  d\theta^{b})=\delta_{ab};\; a,b=1,\dots,\dim\gf-\dim\af$.
%%  
%%  Then for generator of field transformation we can write
%%  \begin{equation} \mathcal{G}_{i}=\left(\frac{2dt}{w_{i}}-\sqrt{\kappa}
%%  d\xi_{t}\right) \partial_{w_{i}}+\frac{\sqrt{\tau}}{w_{i}}\left(\sum_{a=1}^{\dim\gf-\dim\af}\left(d
%%  \theta ^{a} t^{a}_{i}\right)\right)
%%  \label{eq:5}
%%  \end{equation}
%%   Substituting to equation \eqref{eq:8} we get martingale condition
%%  
%%  \begin{equation} \left(-2 \mathcal{L}_{-2}+\frac{1}{2}\kappa
%%  \mathcal{L}_{-1}^{2}+\frac{\tau}{2}\left( \sum_{a=1}^{\dim\gf-\dim\af}
%%  \mathcal{J}^{a}_{-1} \mathcal{J}^{a}_{-1}\right)\right) \mathcal{F}_{\mathbb{H}}=0
%%  \label{eq:9}
%%  \end{equation}
%%  
%%  which can be rewritten as the requirement for
%%  \begin{equation} \psi=\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau
%%  \left(\sum_{a=1}^{\dim\gf-\dim\af}J^{a}_{-1}J^{a}_{-1}\right)\right)
%%  \phi_{(\mu,\nu)}
%%  \label{eq:10}
%%  \end{equation}
%%   to be level two null-field.
%%  
%%   It is not always convenient to work with generators
%%  orthogonal with respect to Killing form $\mathcal{K}$, sometimes it is preferable to use Chevalley
%%  or Cartan-Weyl basis. So we change normalization condition of additional
%%  $\dim\gf$-dimensional Brownian motion \eqref{eq:5} to $\mathbb{E}\left[d\theta^{a}\;
%%  d\theta^{b}\right]=\mathcal{K}(t^{a},t^{b})dt$, and the generator of field transformation
%%  \eqref{eq:2} is
%%  \begin{equation*} \mathcal{G}_{i}=\left(\frac{2dt}{w_{i}}-\sqrt{\kappa}
%%  d\xi_{t}\right) \partial_{w_{i}}+\frac{\sqrt{\tau}}{w_{i}}\left(\sum_{a:\mathcal{K}(t^{a},\tilde{t}^{b})=0}\left(d
%%  \theta ^{a} t^{a}_{i}\right)\right).
%%  \end{equation*}
%%   Martingale condition \eqref{eq:8} can be written as
%%  \begin{equation*} \left(-2 \mathcal{L}_{-2}+\frac{1}{2}\kappa
%%  \mathcal{L}_{-1}^{2}+\frac{\tau}{2}\left( \sum_{a} \mathcal{J}^{a}_{-1} \mathcal{J}^{a}_{-1}-
%%  \sum_{b}\tilde{\mathcal{J}}^{b}_{-1} \tilde{\mathcal{J}}^{b}_{-1}\right)\right)
%%  \mathcal{F}_{\mathbb{H}}=0.
%%  \end{equation*}
%%   This equation can again be rewritten as the requirement for
%%  \begin{equation} \psi=\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau
%%  \left(\sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{-1}-\sum_{b=1}^{\dim\af}\tilde{J}^{b}_{-1}\tilde{J}^{b}_{-1}\right)\right)
%%  \phi_{(\mu,\nu)}
%%  \label{eq:30}
%%  \end{equation}
%%   to be level two null-field.
%%  
%%  Acting with $L_{2}$ we get
%%  \begin{equation*}
%%    L_{2}\psi= \left(-8 L_{0}+c/2+ 3 \kappa L_{0}+\frac{1}{2}\tau k \dim\gf\right)
%%  \phi_{(\mu,\nu)}
%%  \end{equation*}
%%  \begin{equation}
%%    \label{eq:28} (3\kappa-8) h_{(\mu,\nu)}+c+\tau k =0
%%  \end{equation}
%%  
%%  Now we act with $L_{1}^{2}$  and obtain
%%  \begin{equation}
%%    \label{eq:17} 12 h_{(\mu,\nu)}+2\kappa h_{(\mu,\nu)} (2h_{(\mu,\nu)}+1) + \tau
%%  (C_{\mu}-\tilde{C}_{\nu})=0,
%%  \end{equation}
%%   where $C_{\mu}=(\mu,\mu+2\rho)$ is the eigenvalue of quadratic Casimir operator
%%  $\sum_{a}t^{a}t^{a}$. We need to get one more equation since we
%%  have three variables $\kappa,\tau,h_{(\mu,\nu)}$. We can use that $\psi$ is singular vector so
%%  $L_{1}\psi=0$ and then $J_{1}^{a}L_{1}\psi=0,\; \tilde{J}_{1}^{b} L_{1}\psi=0$, so
%%  
%%  \begin{equation}
%%  L_{1} \psi=L_{1}\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau
%%  \left(\sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{-1}-\sum_{b=1}^{\dim\af}\tilde{J}^{b}_{-1}\tilde{J}^{b}_{-1}\right)\right)
%%  \phi_{(\mu,\nu)}
%%  \end{equation}
%%  
%%  \begin{equation*}
%%  L_{1}\psi = \left(-6 L_{-1}+\kappa (2  L_{-1} L_{0}+L_{0}) + \tau \sum_{a=1}^{\dim\gf}J^{a}_{-1}J^{a}_{0}\right)
%%  \phi_{(\mu,\nu)}
%%  \end{equation*}
%%  \begin{equation*}
%%    L_{1}^{2}\psi= \left(-12 L_{0}+\kappa (4  L_{0}^{2}+2 L_{0}) + \tau \sum_{a=1}^{\dim\gf}J^{a}_{0}J^{a}_{0}\right)
%%  \phi_{(\mu,\nu)}
%%  \end{equation*}
%%  \begin{equation*}
%%    J_{1}^{a}L_{1} \psi= \left(-6 J_{0}^{a}+\kappa (2  J_{0}^{a} L_{0}+J_{0}^{a}) + \tau
%%      \sum_{b=1}^{\dim\gf}\sum_{c=1}^{\dim\gf}i f^{abc}J^{c}_{0}J^{b}_{0}+k\tau J_{0}^{a}\right)
%%  \phi_{(\mu,\nu)}
%%  \end{equation*}
%%  
%%  
%%  \begin{equation*}
%%    L_{2}\psi= \left(-8 L_{0}+c/2+ 3 \kappa L_{0}+\frac{1}{2}\tau k \dim\gf\right)
%%  \phi_{(\mu,\nu)}
%%  \end{equation*}
%%  
%%  \begin{eqnarray*}
%%    (3\kappa-8) h_{(\mu,\nu)}+c/2+\frac{1}{2}\tau k \dim\gf=0\\
%%    -12 h_{(\mu,\nu)}+\kappa (4 h_{(\mu,\nu)}^{2}+2 h_{(\mu,\nu)}) +\tau C(\mu) =0\\
%%    (-6 +\kappa (2 h_{(\mu,\nu)}+1)+\tau
%%  \end{eqnarray*}
%%  \begin{equation}
%%    \label{eq:29} \left(-6 + \kappa + \tau + 2 \kappa h_{(\mu,\nu)}\right) J^{3}_{0}
%%  \phi_{(\mu,\nu)}=0
%%  \end{equation}
%%  
%%  Acting on \eqref{eq:30} with $L_{2}$ and $L_{1}^{2}$ we can obtain equations \eqref{eq:28} and
%%  \eqref{eq:17}. To get more equations we can act by $\afh$-generators $\tilde{J}^{b}_{1}$ and $L_{1}$
%%  similarly to what was done with $J^{3}_{1}$ to get equation \eqref{eq:29}.
%%  
%%  To obtain simpler relations it is also possible to use generalization of Knizhnik-Zamolodchikov
%%  equations \eqref{eq:6} as was done for WZNW-models in paper \cite{alekseev2010sle}.
%%  
%%  
%%  Now consider simple example. Let $G=SU(2)$ and $A=U(1)$, and corresponding Lie algebras
%%  $\gf=\mathrm{su}(2)$ with generators $J^{1},J^{2},J^{3}$ and $\af=\mathrm{u}(1)$ with the generator
%%  $J^{3}$, $\af\subset\gf$. Note that $\mathcal{K}(J^{a},J^{b})=2\delta^{ab}$.
%%  
%%   The equation \eqref{eq:10} is now
%%  \begin{equation}
%%    \label{eq:11} \psi=\left(-2L_{-2}+\frac{1}{2}\kappa L_{-1}^{2}+\frac{1}{2}\tau
%%  \left(J^{1}_{-1}J^{1}_{-1}+J^{2}_{-1}J^{2}_{-1}\right)\right) \phi_{(\mu,\nu)}
%%  \end{equation}
%%   If we use basis $J^{+}=\frac{J^{1}+iJ^{2}}{\sqrt{2}},\;
%%  J^{-}=\frac{J^{1}-iJ^{2}}{\sqrt{2}}$ this equation is rewritten in the form
%%  \begin{equation} \psi= \left(-2
%%  L_{-2}+\frac{\kappa}{2}L_{-1}^{2}+\frac{\tau}{2}\left[J^{+}_{-1}J^{-}_{-1}+J^{-}_{-1}J^{+}_{-1}\right]\right)
%%  \phi_{(\mu,\nu)}
%%  \label{eq:12}
%%  \end{equation}
%%   which is similar to the equations for parafermionic fields introduced in paper
%%  \cite{santachiara2008sle}.
%%  
%%  Central charge in this case is equal to
%%  \begin{equation}
%%    \label{eq:14} c=\frac{2(k-1)}{k+2}
%%  \end{equation}
%%  
%%  Let us study the solutions of the equation \eqref{eq:11}. Act on $\psi$ with $L_{2}$:
%%  \begin{equation}
%%    \label{eq:13} L_{2}\psi=(-8L_{0}+c+3\kappa L_{0}+\tau k)\phi_{(\mu,\nu)}=0,
%%  \end{equation}
%%   and obtain
%%  \begin{equation}
%%    \label{eq:28} (3\kappa-8) h_{(\mu,\nu)}+c+\tau k =0
%%  \end{equation}
%%   Another equation can be obtained by action of $L_{1}^{2}$
%%  \begin{equation}
%%    \label{eq:15} L_{1}^{2}\psi = (12 L_{0} + \kappa(4 L_{0}^{2}+2 L_{0}) +\tau
%%  (J_{0}^{1}J_{0}^{1}+J_{0}^{2}J_{0}^{2}))\phi_{(\mu,\nu)}=0
%%  \end{equation}
%%   Since
%%  $J_{0}^{1}J_{0}^{1}+J_{0}^{2}J_{0}^{2}=J_{0}^{1}J_{0}^{1}+J_{0}^{2}J_{0}^{2}+J_{0}^{3}J_{0}^{3}-J_{0}^{3}J_{0}^{3}$
%%  we have
%%  \begin{equation}
%%    \label{eq:17} 12 h_{(\mu,\nu)}+2\kappa h_{(\mu,\nu)} (2h_{(\mu,\nu)}+1) + \tau
%%  (C_{\mu}-\tilde{C}_{\nu})=0,
%%  \end{equation}
%%   where $C_{\mu}=(\mu,\mu+2\rho)$ is the eigenvalue of quadratic Casimir operator
%%  $\sum_{a}t^{a}t^{a}$. For $u(1)$ it is just $\nu^{2}$. We need to get one more equation since we
%%  have three variables $\kappa,\tau,h_{(\mu,\nu)}$. We can use that $\psi$ is singular vector so
%%  $L_{1}\psi=0$ and then $J_{1}^{3}L_{1}\psi=0$, so
%%  \begin{equation}
%%    \label{eq:29} \left(-6 + \kappa + \tau + 2 \kappa h_{(\mu,\nu)}\right) J^{3}_{0}
%%  \phi_{(\mu,\nu)}=0
%%  \end{equation}
%%  
%%  
%%  Turn to the case $k=2, \;c=1/2$ which corresponds to square lattice Ising model. We have three
%%  non-equivalent fields numbered by the pairs of $su(2)$ and $u(1)$ weights, which are $\phi_{(0,0)},
%%  \; h_{(0,0)}=0; \quad \phi_{(0,2)}, \; h_{(0,2)}=1/2; \quad \phi_{(1,1)}, \; h_{(1,1)}=1/16$. The
%%  equation \eqref{eq:29} is trivial for $\phi_{(0,0)}$ and $\phi_{(0,2)}$, so the equations
%%  \eqref{eq:17}, \eqref{eq:28} are consistent although we do not know how to interpret $\tau$ in
%%  context of the Ising model. If we consider the field $\phi_{(1,1)}$ we have no solution for
%%  $\kappa,\tau$, so we see that not every primary field corresponds to some boundary condition
%%  changing operator.
%%  
%%  
%%  \section{Algebraic equations}
%%  \label{sec:algebraic-equations}
%%  
%%  \section{Examples}
%%  \label{sec:examples}
%%

\bibliography{bibliography}{}
\bibliographystyle{utphys}

\end{document}