### 2012-08-16 AffineLieAlgebras/Thesis/concl.tex
Some notes for conclusions

\conclusion
\label{cha:conclusion}

В работе впервые получены следующие результаты

%%
%% End of file
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: