### 2012-11-17 AffineLieAlgebras/MQFT2012/paper.tex
Started the paper

\documentclass[12pt]{article}
\usepackage{amsfonts}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath,amssymb,amsthm}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[english]{babel}

\newtheorem{Def}{Definition}[section]
\newtheorem{theorem}{Theorem}
\newtheorem{statement}{Statement}
\newtheorem{Cnj}[Def]{Conjecture}
\newtheorem{Prop}[Def]{Property}
\newtheorem{example}{Example}[section]

\newcommand{\co}[1]{\stackrel{\circ }{#1}}
\newcommand{\gf}{\mathfrak{g}}
\newcommand{\nfp}{\mathfrak{n}^{+}}
\newcommand{\nfm}{\mathfrak{n}^{-}}
\newcommand{\af}{\mathfrak{a}}
\newcommand{\uf}{\mathfrak{u}}
\newcommand{\sfr}{\mathfrak{s}}
\newcommand{\aft}{\widetilde{\mathfrak{a}}}
\newcommand{\afb}{\mathfrak{a}_{\bot}}
\newcommand{\hf}{\mathfrak{h}}
\newcommand{\hfb}{\mathfrak{h}_{\bot}}
\newcommand{\pf}{\mathfrak{p}}

\newcommand{\gfh}{\hat{\mathfrak{g}}}
\newcommand{\afh}{\hat{\mathfrak{a}}}
\newcommand{\sfh}{\hat{\mathfrak{s}}}
\newcommand{\bff}{\mathfrak{b}}
\newcommand{\hfg}{\hf_{\gf}}

\begin{document}
\title{On singular elements in conformal field theory. \\ Singular elements in vertex operator algebras}

\author{V.D.~Lyakhovsky$^1$, A.A.~Nazarov$^{1,2}$ \\
  {\small $^1$ Department of High-energy and elementary particle physics, SPb State University}\\
  {\small 198904, Saint-Petersburg, Russia,}
  {\small e-mail: lyakh1507@nm.ru}\\
  {\small$^{2}$ Chebyshev Laboratory,}
  {\small Department of Mathematics and Mechanics, SPb State University}\\
  {\small 199178, Saint-Petersburg, Russia}
  {\small email: antonnaz@gmail.com}}
\date{}
\maketitle

\begin{abstract}
  We consider singular elements of simple finite-dimensional and affine Lie algebra modules. We
  construct corresponding vertex operator algebras. We consider decompositions of singular elements
  connected with branching, splints and parabolic Verma modules and describe vertex counterparts. 
\end{abstract}

\section{Introduction}
The notion of vertex operator algebra \cite{borcherds1986vertex} formalizes
algebraic structure that underlies two-dimensional conformal field theory
\cite{kac1998vertex,frenkel1988vertex}. Vertex algebras associated to
representations of affine Lie algebras \cite{frenkel1992vertex} form an
important class corresponding to Wess-Zumino-Novikov-Witten
\cite{witten1984nab,Walton:1999xc} and coset models \cite{Goddard198588} of
conformal field theory.

We show that recurrent relations on branching coefficients for affine Lie
algebra modules are related to the structure of Virasoro modules.

We study singular elements of affine Lie algebra modules and establish the properties
of

\begin{itemize}
\item Singular element of simple Lie algebra irreducible module, VOA corresponding to that module.
  What are the relations in VOA on singular vectors? Is there BGG-resolution?
\item Affine Lie algebra, parabolic Verma modules?
\item Splints?
\item Coset vertex operator algebras and branching
\end{itemize}

\begin{Def}
  Vertex algebra: $V$ -- vector space with identity $1$, endomorphism
  $T:V\to V$ and linear multiplication map $Y:V\otimes V\to V((z)):
  (a,b)\to Y(a,z)b=\sum_n a_n b z^{-n-1}$,
  such that
  \begin{itemize}
  \item {\it Identity} $\forall a\in V, Y(1,z) a =a$ and $Y(a,z) 1 \in a+z V z$
  \item {\it Translation} $T(1)=0$ and $\forall a,b\in V$:
    $$Y(a,z)T b - T Y(a,z) b =
    \frac{d}{dz} Y(a,z) b$$
  \item {\it Four point function}. $\forall a,b,c\in V \exists
    X(a,b,c;z,w)\in V[[z,w]][z^{-1},w^{-1},(z-w)^{-1}]: Y(a,z)Y(b,w)c,
    Y(b,w)Y(a,z)c, Y(Y(a,z-w)b,w)c$ are expansions of $X(a,b,c;z,w)$
    in $V((z))((w)), V((w))((z)), V((w))((z-w))$
  \item {\it Grading} $V=\bigoplus_{n=0}^{\infty} V_n$
  \item {\it Virasoro element} $\omega\in V_2: Y(w,z)=\sum_{n\in
      \mathbb{Z}} L_n z^{-n-2}$ satisfies $\forall a\in V_n$ the
      relations $L_0 a =n a$,
      $Y(L_{-1}a,z)=\frac{d}{dz}Y(a,z)=[Y(a,z),T]$,
      $[L_m,L_n]a=(m-n)L_{m+n} a + \delta_{m+n,0} \frac{m^3 -
        m}{12}ca$, $c$ -- central charge.
  \end{itemize}

\end{Def}

\bibliography{bibliography}{}
\bibliographystyle{utphys}

\end{document}