### 2013-02-07 AffineLieAlgebras/Simons/talk.tex
Created folder for the talk at Simons Center

\documentclass[a4paper]{article}

\begin{document}
Hello!
It's a great pleasure for me to be here in Simons Center for geometry and physics. 
My talk is called 
\end{document}