### 2013-02-09 AffineLieAlgebras/Simons/talk.tex
Some work on the talk

\documentclass[a4paper]{article}

\title{Boundary condition changing operators and singular vectors of Virasoro and
affine Lie algebra modules}

\author{Anton Nazarov}
\begin{document}
\maketitle
\begin{abstract}
We consider the behaviour of CFT correlation functions with respect to
Schramm-Loewner evolution with additional Brownian motion on gauge
group. Using the analogy with lattice models we consider
correlation functions that satisfy martingale conditions and contain
boundary condition changing operators and primary fields.
Martingale conditions can be rewritten as algebraic relations on
boundary condition changing operators and interpreted from
representation-theoretic point of view.
\end{abstract}

Hello!
It's a great pleasure for me to be here in Simons Center for geometry and physics. 
My talk is called ``Boundary condition changing operators and singular vectors of Virasoro and
affine Lie algebra modules''.

Consider some simple two-dimensional lattice model, for example, Ising
model on hexagonal lattice. Let us study the behavior of domain walls.
In order to do it we need to look at the model on the bounded domain.
We take upper half-plane. Then impose boundary condition in such a way
that we have a domain wall from zero to infinity.

In critical model there is continuous limit. It is given by a solution
of Schramm-Loewner stochastic differential equation. 
\end{document}