### 2013-05-06 AffineLieAlgebras/Simons/tricritical-ising-sle.tex
Some notes for tricritical Ising

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm,amsfonts}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}

\newcommand{\co}[1]{\stackrel{\circ }{#1}}
\newcommand{\gf}{\mathfrak{g}}
\newcommand{\nfp}{\mathfrak{n}^{+}}
\newcommand{\nfm}{\mathfrak{n}^{-}}
\newcommand{\af}{\mathfrak{a}}
\newcommand{\uf}{\mathfrak{u}}
\newcommand{\sfr}{\mathfrak{s}}
\newcommand{\aft}{\widetilde{\mathfrak{a}}}
\newcommand{\afb}{\mathfrak{a}_{\bot}}
\newcommand{\hf}{\mathfrak{h}}
\newcommand{\hfb}{\mathfrak{h}_{\bot}}
\newcommand{\pf}{\mathfrak{p}}

\newcommand{\gfh}{\hat{\mathfrak{g}}}
\newcommand{\afh}{\hat{\mathfrak{a}}}
\newcommand{\sfh}{\hat{\mathfrak{s}}}
\newcommand{\bff}{\mathfrak{b}}
\newcommand{\hfg}{\hf_{\gf}}

\begin{document}
\title{Schramm-Loewner evolution in tricritical Ising model}

\maketitle

Tricritical Ising: integrability (RSOS), supersymmetry (?), conformally-invariant boundary conditions, identification with CFT. 
CFT with supersymmetry (additional generators and relations). 
Boundary conditions, CFT primary fields and SLE. 
SLE on superspace \cite{nagi2005stochastic,rasmussen2004stochastic} and BCFT for tricritical Ising (parameter correspondence).

Proof: discrete holomorphicity and integrability. 
(Loop description?)
\bibliography{bibliography}{} 
\bibliographystyle{utphys}

\end{document}