### 2014-02-01 area-length/notes.tex
Some text on Liouville and triangulations

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm,amsfonts}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\newcommand{\co}[1]{\stackrel{\circ }{#1}}
\newcommand{\gf}{\mathfrak{g}}
\newcommand{\nfp}{\mathfrak{n}^{+}}
\newcommand{\nfm}{\mathfrak{n}^{-}}
\newcommand{\af}{\mathfrak{a}}
\newcommand{\uf}{\mathfrak{u}}
\newcommand{\sfr}{\mathfrak{s}}
\newcommand{\aft}{\widetilde{\mathfrak{a}}}
\newcommand{\afb}{\mathfrak{a}_{\bot}}
\newcommand{\hf}{\mathfrak{h}}
\newcommand{\hfb}{\mathfrak{h}_{\bot}}
\newcommand{\pf}{\mathfrak{p}}

\newcommand{\gfh}{\hat{\mathfrak{g}}}
\newcommand{\afh}{\hat{\mathfrak{a}}}
\newcommand{\sfh}{\hat{\mathfrak{s}}}
\newcommand{\bff}{\mathfrak{b}}
\newcommand{\hfg}{\hf_{\gf}}

\begin{document}
\title{Notes on area-length problem for CLE}
%\author{Anton Nazarov}%% $^{1,2}$}

\maketitle

\begin{abstract}
%  Here I document staff related to my work on area-length problem.

  We consider conformal loop ensemble. The ratio
  of loop area to certain power of loop length exhibit universal
  scaling behavior. Scaling function is proportional to Bessel's
  function. We show it using boundary Liouville field theory fusion rules for degenerate field.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

One can ask simple question: how many loops is there?

To make this question more precise consider conformal loop ensemble \cite{sheffield2010conformal}. It is critical limit of
lattice different models for different values of $\kappa$. Loops can be interpreted as a critical
limits of domain walls in lattice models. 
We are interested in distribution of loops by length and area.

Though several publications exist
\cite{cardy2003exact,cardy2003crossover,cardy2001exact,cardy1994geometrical}
on this problem and the answer is
obtained from computer simulation \cite{ richard2001scaling}, good understanding was not
obtained.

Some results exist in literature for particular case of self-avoiding walks ($\kappa=\frac{8}{3}$). We start
with the description of these results \ref{sec:self-avoid-polyg}, then show that general problem
leads to Liouville field theory \ref{sec:scaling-random-loops} and obtain our result using Liouville
field theory fusion rules \ref{sec:liouville-theory}.

\subsection{Self-avoiding polygons}
\label{sec:self-avoid-polyg}

The model of self-avoiding polygons was numerically studied in the paper \cite{richard2001scaling}.
The simplest case is the generating function for polygons of given length. It can be introduced as
\begin{equation}
  \label{eq:86}
  G(x)=\sum_l p_l x^l,
\end{equation}
where  $p_l$ is the number of different polygons on square lattice of length $l$. Different are
polygons that can not be superposed by a lattice translation.
If we know generating function we can obtain number of polygons of given length differentiating
w.r.t. $x$.

We can also introduce generating function for polygons of given length and area:
\begin{equation}
  \label{eq:75}
  G(x,q)=\sum_{m,n} p_{m,n} x^{m} q^{n},
\end{equation}
where $p_{m,n}$ is the number of polygons with perimeter $m$, enclosing an area $n$ passing through
a given edge (or node). 
Cardy \cite{cardy2001exact} use following definition for generating function of rooted loops, using
$g=-\ln q$:
\begin{equation}
  \label{eq:76}
  G^{(r)}(x,g)=\sum_{l,A} p_{l,A}^{(r)} x^{l} e^{-gA}
\end{equation}
Generating function for rooted polygons $G^{(r)}(x,g)$ is related to $G(x,g)$ by
\begin{equation}
  \label{eq:81}
  G^{(r)}(x,g)=x \frac{d}{dx} G(x,g),
\end{equation}
because root can be chosen in $l$ ways.

This model has critical point $(x_{c},0)$ and generating functions displays scaling behavior
\begin{equation}
  \label{eq:77}
  G(x,g)\sim G(x,g)_{\mathrm{reg}}+g^{\theta} F\left((x_{c}-x)g^{-\phi}\right)
\end{equation}
Critical indices $\theta,\phi$ are related to conventional exponents $\nu,\alpha$:
\begin{eqnarray}
  \label{eq:78}
  \phi=\frac{1}{2}\nu\\
  \theta=\frac{1}{2}(1-\alpha) \nu
\end{eqnarray}
Hyperscaling hypothesis adds another relation
\begin{equation}
  \label{eq:79}
  \alpha=2-2\nu
\end{equation}
Numerical results predict
\begin{eqnarray}
  \label{eq:80}
  \nu=\frac{3}{4}\\
  \theta=\frac{1}{3}\\
  \phi=\frac{2}{3}
\end{eqnarray}
We want to find scaling function $F$.

\subsection{$O(n)$ model}
\label{sec:introduction-1}

Self-avoiding polygons can be thought as the limit $n\to 0$ of loops in $O(n)$ model.

Partition function for $O(n)$ model is
\begin{equation}
  \label{eq:9}
  Z_{O(n)} = \sum_{\mbox{loop conf}} x^{\mbox{total length}}
  n^{\mbox{number of loops}}
\end{equation}
Critical point is $x=\frac{1}{\sqrt{2+\sqrt{2-n}}}$, the continuum
limit in this point is described by minimal model $(p,p')$ with
$n=-2\cos\left(\frac{\pi p}{p'}\right)$. $c=1-6\frac{(p-p')^2}{pp'}$, $b=\sqrt{\frac{3}{2}}$.

Self-avoiding walk is obtained in the limit $n\to 0$. Then $c=0$ and
we have logarithmic conformal field theory.

Logarithmic CFT $\sim$ Liouville theory ???

\section{Scaling of random loops}
\label{sec:scaling-random-loops}

Cardy \cite{cardy2001exact} argues that generating function $G^{(r)}$ for rooted self-avoiding loops is given by a derivative w.r.t. $x$
of expectation value of $e^{-gA}$ (eq.6):
\begin{equation}
  \label{eq:82}
  Z=\left<e^{-gA}\right>_{\mathrm{SAL}}=\left<e^{-\sqrt{g}\int J_{\lambda}A_{\lambda}d^{2}x}\right>_{\mathrm{SAL}.A}
\end{equation}
But reasoning is quite strange and unsound. Due to the need to use $g$ for metric in later sections
we will denote area coupling constant by $\mu$ and loop length coupling constant by $\mu_{B}=-\ln x$
(boundary).

In this section we introduce scaling function for random loops.

Consider rooted loops in $O(n)$ model. Let $N^r(L)$ be number of
rooted loops with a given length.  By $N^r(L,A)$ we denote the number
of loops of given length and area. $N^r(L)=\int_0^{\infty} N^r(L,A)
dA$. Area weighted number of loops is denoted by $G^r(L,\mu)$. (Here
$\mu$ is bulk coupling constant ).
\begin{equation}
  \label{eq:2}
  G^r(L,\mu)=\int e^{\mu A} N^r(L,A) dA
\end{equation}

\begin{equation}
  \label{eq:1}
  G^r (\mu,x) = \sum_{\mathrm{clusters}} N^r (L,A) x^L e^{-\mu A}
\end{equation}

The idea is to write $G$ as a correlation function in Liouville
theory.

Note that number of rooted loops of given area scales as the area to
the power minus one.

Define Hausdorff dimension as follows.
\begin{definition}
  {\it Hausdorff content} of the set $S$ is
  \begin{equation}
    \label{eq:3}
    C^d_H = \mathrm{inf} \sum_i r_i^d,
  \end{equation}
  where $\{r_i\}$ are such that there exists cover of $S$ by balls of
  radii $r_i>0$.

  {\it Hausdorff dimension}
  \begin{equation}
    \label{eq:4}
    d_H\equiv\mathrm{dim}_H(S) = \mathrm{inf}\{ d\geq 0: C^d_H(S)=0\}
  \end{equation}
\end{definition}

Consider loop of length $L$. The radius of the loop is connected with
the length as $L\sim R^{d_H}$. The area is related to radius $A\sim
R^2$, so $A\sim L^{\frac{2}{d_H}}$.
(It is proved or almost proved in papers by Lawler).

For CLE the connection of Hausdorff dimension and parameter $\kappa$
is known:
\begin{equation}
  \label{eq:5}
  d_H=1+\frac{\kappa}{8}
\end{equation}

($\nu^{-1}=d_H$)

At $x\to x_c$
\begin{equation}
  \label{eq:6}
  G^r(\mu,x)= \mu^{\theta} F\left(\frac{x-x_c}{\mu^{\phi}}\right)
\end{equation}

Partition function is product over loops of different length
\begin{equation}
  \label{eq:10}
  Z_{O(n)} = \sum_{\mbox{loop conf}} \prod_L x^{L N(L)} n^{N(L)} = \sum_{\mbox{loop conf}} \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)}
\end{equation}

We can rewrite $G^r(\mu,x)$ as the derivative of generating function
\begin{equation}
  \label{eq:7}
  G^r(\mu,x)=\mu\partial_x \log Z(\mu,\mu_B)
\end{equation}

Here
\begin{equation}
  \label{eq:8}
  Z(\mu,\mu_B) = \left< e^{-\mu A} \right> = \frac{\sum_{\mbox{loop conf}}
  \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)} e^{-\mu A}}{\sum_{\mbox{loop conf}} \prod_L\prod_A x^{L N(L,A)} n^{N(L,A)}}
\end{equation}

\section{Liouville theory}
\label{sec:liouville-theory}

Liouville theory, flat metric, asymptotic for field, Coulomb gas formalism.
\cite{nakayama2004liouville,ponsot2002boundary,teschner2001liouville,fateev2000boundary,teschner2000remarks,zamolodchikov1996conformal}

Liouville theory describes internal metric of surface. It is possible to obtain length of boundary
and area of surface with such a metric. To get the answer for flat loops we need to impose some
condition on metric in the bulk.

The action for Liouville theory on a surface $\Gamma$ in arbitrary background metric $g$ is
\begin{equation}
  \label{eq:13}
  S=\frac{1}{4\pi} \int_{\Gamma} \sqrt{g} \left( g^{ab} \partial_{a}\varphi \partial_{b}\varphi + Q R \varphi +4\pi \mu e^{2b\varphi}\right) d^{2}x
\end{equation}

This theory gives internal metric $\tilde g^{ab}=e^{\varphi(x)}g^{ab}(x)$

Classical equations of motion are
\begin{equation}
  \label{eq:16}
  \Delta \varphi(x) = \frac{1}{2} \left(b+\frac{1}{b}\right) R (x) + 4\pi \mu e^{2 b \varphi}
\end{equation}
Here $\Delta$ is Laplace-Beltrami operator
\begin{equation}
  \label{eq:17}
  \Delta=\frac{1}{\sqrt{g}}\partial_{\mu} \left(\sqrt{g} g^{\mu\nu} \partial_{\nu}\right)
\end{equation}

Since
\begin{equation}
  \label{eq:93}
  \sqrt{g}(R-\Delta \varphi)=\sqrt{\tilde g} \tilde R, 
\end{equation}
for metric $\tilde g= e^{2b\varphi} g$, equation of motion can be written as
\begin{equation}
  \label{eq:94}
  \tilde R(x) + 4\pi \mu=0
\end{equation}
So classical configurations have constant curvature $-4\pi\mu$.

In most cases it is possible to use flat background metric, but one needs to be careful with
singularities. For example, consider a manifold with the topology of a sphere. We can use flat
background metric but we need to introduce singularity at infinity \cite{zamolodchikovlectures}.

In flat background metric $g: ds^{2}=dz d\bar z$ equations of motions are
\begin{equation}
  \label{eq:90}
  4\partial\bar \partial \varphi = 4\pi\mu e^{2b\varphi}
\end{equation}
To solve it we consider auxiliary equation:
\begin{equation}
  \label{eq:91}
  4\partial^{2} \psi + t\psi=0
\end{equation}
Function $t$ here is $zz$ component of energy-momentum tensor. Denote by $\psi$  vector
$\psi=(\psi_{1},\psi_{2})$ of linearly independent solutions. Then solution of equation
\eqref{eq:90} is given by the expression
\begin{equation}
  \label{eq:92}
  \varphi=-\frac{1}{b} \log \left(\bar \psi(\bar z) \lambda \psi(z)\right),
\end{equation}
where $\lambda$ is some matrix that is invariant w.r.t. to the action of monodromy group around the
singularity.

\begin{equation}
  \label{eq:89}
  g=
\end{equation}

In theory with the action
$$S=\frac{1}{4\pi} \int \sqrt{g}\left( \frac{1}{2} g^{\mu\nu} \partial_{\mu} \varphi \partial_{\nu}
  \varphi + Q R\varphi + 4\pi\mu e^{2b\varphi}\right) d^{2}x,$$ 
where $g$ is a background metric we need to assume some regularization procedure to define a
functional integral, because we want to integrate over measure that does not depend upon metric.
The idea is that the form of the action is the same but parameters are changed. 
So now $\mu$ is not a classical cosmological constant and $\mu,b,Q$ are connected through
consistency requirements for the theory. 
The consistency requirement is background-independence. The action should not change when we choose
another background metric and shift field $\varphi$ accordingly. 
Let $\tilde g = e^{\sigma} g$, redefine $\varphi$ as $\tilde \varphi=\varphi + \frac{Q}{2}\sigma$,
then all the terms with $\sigma$ are cancelled if we take into account contributions from matter
partition function. But it is possible only if 
$$Q=\frac{25-c_{M}}{6}$$. 
To make exponential term background-independent we need to have $Q=\frac{1}{b}+b$.

If we consider correlation functions of fields $V_{\alpha}=e^{2\alpha\varphi}$ and impose
background-independence conditions we obtain relation on conformal weights
$$\Delta_{\alpha}=\alpha(Q-\alpha)$$
So for theory to be background-independent we need to have these relations.

Conformal transformation $z\to w(z)$ can be presented as a change in background metric, so 
$$\varphi(w,\bar w) = \varphi(z,\bar z) -\frac{Q}{2}\log \left|\frac{dw}{dz}\right|^{2}$$
If we consider flat background metric, we need to have asymptotic $\varphi(z,\bar z)\sim -Q \log
|z|^{2}$ at infinity to have smooth metric $e^{2b\varphi}$.

Now if we consider flat theory with $\mu=0$ everything is good, but when we introduce the observable
$e^{-\mu A}=e^{-\mu \int d^{2}x e^{2b\varphi}}$ we need to regularize exponential field, impose
consistency conditions and we get ordinary Liouville theory as the result! 
So Liouville theory seems to be unique in this way.

My understanding is that averaging over flat configurations weighted by $e^{-\mu A}$ is equivalent
to averaging over configurations with cosmological constant $\mu$ which introduces non-zero
curvature of singularity at infinity in flat background.

By the way, it is also written that they consider unnormalized correlation functions (without the
division by $Z$) because there are difficulties with the definition of functional integral $\int
D\varphi e^{-S[\varphi]}$. 
So we need to impose some normalization on their results.

We can select locally flat metric $g^{ab}=\delta^{ab}$ on some sheet, then the action takes the form
\begin{equation}
  \label{eq:14}
  S= \int d^{2}x \frac{1}{4\pi} (\partial_{a} \varphi)^{2}+\mu e^{2b\varphi}
\end{equation}
Since we rewrote the action in this form we need to take into account the curvature. We do it by
supposing the existence of conical singularity at infinity.
\begin{equation}
  \label{eq:15}
  \varphi(z)=-Q \log (z\bar z)+O(1) \quad z\to\infty
\end{equation}
The equations of motion take the from
\begin{equation}
  \label{eq:18}
  \Delta \varphi = 4\pi b e^{2b\varphi}
\end{equation}
The background charge is connected with the parameter $b$ by equation
\begin{equation}
  \label{eq:20}
  Q=b+\frac{1}{b}
\end{equation}

Liouville field theory is a conformal field theory with central charge
\begin{equation}
  \label{eq:19}
  c=1+6\left(b+\frac{1}{b}\right)^{2}
\end{equation}
and primary fields are given by vertex operators $V_{\alpha}=e^{\alpha\varphi}$. But primary fields
$V_{\alpha}$ and $V_{Q-\alpha}$ are to be identified. 
Reflection amplitudes are used for this identification:
\begin{equation}
  \label{eq:87}
  V_{\alpha}(z,\bar z)=S(\alpha) V_{Q-\alpha}(z,\bar z),
\end{equation}
where
\begin{equation}
  \label{eq:88}
  S(\alpha)=\frac{\left(\pi\mu\gamma(b^{2})\right)^{\frac{Q-2\alpha}{b}}}{b^{2}}
  \frac{\gamma(2\alpha b -b^{2})}{\gamma\left(2-\frac{2\alpha}{b}+\frac{1}{b^{2}}\right)}
\end{equation}

For $\alpha=\frac{Q}{2}+i P, P\in \mathbb{R}$
fields $V_{\alpha}$ are non-local. Conformal dimension of primary field $V_{\alpha}$ is
\begin{equation}
  \label{eq:32}
  \Delta_{\alpha}=\alpha(Q-\alpha)
\end{equation}

Fields $V_{-\frac{nb}{2}}, \; n\in \mathbb{N}$ are degenerate, correlation functions of these fields
satisfy differential equations of order $n+1$.

Computation of correlation functions in Liouville field theory relies on screening operators and
recurrent relations that appear from OPE for auxiliary correlation functions.
\subsection{Coulomb gas representation and screening operators}
\label{sec:coul-gas-repr}

Calculation of correlation functions in Liouville field theory can be done as follows. 
Consider correlation some function with operator $V_{-\frac{b}{2}}$, it satisfies second order
differential equation:
\begin{equation}
  \label{eq:33}
  \left(\frac{1}{b^{2}}\partial^{2}+T(z)\right)\left<V_{-\frac{b}{2}}(z) \dots \right>=0
\end{equation}
Now to find some correlation function $\left<V_{\alpha_{1}}(z_{1})\dots
  V_{\alpha_{k}}(z_{k})\right>$ we consider auxiliary correlation function with additional
operator $V_{-\frac{b}{2}}(z)$, find a solution to differential equation and fix the integration
constants using fusion rules for field $V_{-\frac{b}{2}}$.

It can be shown (add reference here? \cite{nakayama2004liouville},\cite{teschner2001liouville}) that 
\begin{equation}
  \label{eq:34}
  V_{-\frac{b}{2}}\times V_{\alpha}= [V_{\alpha-\frac{b}{2}}]+[V_{\alpha+\frac{b}{2}}]
\end{equation}

Full form of the operator product expansion is
\begin{equation}
  \label{eq:39}
  A(z)B(w)\underset{z\to w}{\longrightarrow} \sum_{C} \frac{f_{AB}^{C}}{(z-w)^{h_{A}+h_{B}-h_{C}}} C(w)
\end{equation}

We use auxiliary correlation function that can be computed using Coulomb gas picture. The
computation relies on following ideas.

\subsubsection{Coulomb gas picture. Free boson theory. Minimal models}
\label{sec:free-boson-minimal-models}

\begin{equation}
  \label{eq:40}
  S=\int d^{2}z b (\partial \varphi)^{2},\quad \left<\varphi(z,\bar z)\varphi(w,\bar
    w)\right>=-\frac{1}{2b}\left(\ln (z-w) + \ln (\bar z - \bar w)\right)+const
\end{equation}

\begin{equation}
  \label{eq:42}
  T(z)=-\frac{1}{2}:\partial \varphi \partial \varphi:, \quad c=1
\end{equation}
In holomorphic sector primary fields are $\partial \varphi$ with $h=1$ and
$V_{\alpha}(z)=e^{i \alpha \varphi(z)}$ with dimension $h_{\alpha}=\frac{\alpha^{2}}{2}$.

Field $\varphi(z)$ is not purely holomorphic, it contains zero mode
\begin{equation}
  \label{eq:43}
  \varphi(z)=\varphi_{0}-ia_{0}\ln z+\sum_{n\neq 0}\frac{1}{n}a_{n} z^{-n},\quad
  [a_{n},a_{m}]=n\delta_{n+m,0}, \quad [\varphi_{0},a_{0}]=i
\end{equation}
But we can limit ourselves to chiral operators.

If $\alpha_{1}+\dots+\alpha_{n}=0$ then 
\begin{equation}
  \label{eq:41}
  \left<e^{\alpha_{1}\varphi}(z_{1})\dots e^{\alpha_{n}\varphi}(z_{n})\right> = \prod_{i<j}(z_{i}-z_{j})^{-\alpha_{i}\alpha_{j}}
\end{equation}
Since $\left< :e^{A_{1}}::e^{A_{2}}:\dots :A^{A_{n}}:\right> =
\exp\sum_{i<j}\left<A_{i}A_{j}\right>$, we can take $A_{i}=\alpha_{i}\varphi(z_{i})$ and use the
fact that $\left<A_{i}A_{j}\right>=\ln |z_{i}-z_{j}|^{-\alpha_{i}\alpha_{j}}$.

Neutrality condition can be seen from the invariance of the action w.r.t. $\varphi\to\varphi+a$,
then the correlator obtains factor $e^{\sum \alpha_{i}}$ which is 1 if $\sum \alpha_{i}=0$.

Vacuum expectation value of operator with non-zero charge $\alpha$ should be zero.

To obtain models with $c\neq 1$ one needs to introduce the background charge:
\begin{equation}
  \label{eq:44}
  S=\frac{1}{8\pi} \int d^{2}x \sqrt{g} (\partial_{\mu}\varphi\partial^{\mu}\varphi+2\gamma\varphi R)
\end{equation}
Here $R$ is scalar curvature, $\gamma$ - constant.

(Gauss-Bonnet theorem: $\int d^{2}x \sqrt{g} R = 8\pi(1-h)$, $h$-number of handles).

After introducing interaction with the curvature Ward identities are changed and now neutrality
condition is $\sum \alpha_{i}=2\alpha_{0}$, where $i\sqrt{2}\alpha_{0}=\gamma$. 
Stress-energy tensor now takes the form
\begin{equation}
  \label{eq:45}
  T_{\mu\nu}=T^{(0)}_{\mu\nu}-\frac{\gamma}{2\pi}\left(\partial_{\mu}\partial_{\nu}\varphi-\frac{1}{2}\eta_{\mu\nu}\partial^{\rho}\partial_{\rho}\varphi\right)
\end{equation}
and
\begin{equation}
  \label{eq:46}
  T(z)=-\frac{1}{2}:\partial \varphi\partial\varphi:+i\sqrt{2}\alpha_{0}\partial^{2}\varphi
\end{equation}
If we calculate OPE of $T$ we see that $\partial\varphi$ is not primary any more \cite{difrancesco1997cft}:
\begin{equation}
  \label{eq:47}
  T(z)\partial\varphi(w)\sim
  \frac{i2\sqrt{2}\alpha_{0}}{(z-w)^{3}}+\frac{\partial\varphi(w)}{(z-w)^{2}} +\frac{\partial^{2}\varphi(w)}{z-w}
\end{equation}
Vertex operators are still primary, but conformal dimensions are now
\begin{equation}
  \label{eq:48}
  h_{\alpha}=\alpha^{2}-2\alpha_{0}\alpha
\end{equation}
and central charge is
\begin{equation}
  \label{eq:49}
  c=1-24\alpha_{0}^{2}
\end{equation}
The correlator \eqref{eq:41} is preserved. Now vertex operators $V_{\alpha}$ and
$V_{2\alpha_{0}-\alpha}$ are equivalent, so two-point function
$\left<V_{\alpha}(z)V_{\alpha}(w)\right>$  is not zero, but it should be represented as
$\left<V_{\alpha}(z)V_{2\alpha_{0}-\alpha}(w)\right>=\frac{1}{(z-w)^{h_{\alpha}}}$.

In general we insert non-local screening operator with zero conformal dimension but nonzero charge.
If $\psi$ is primary field with $h_{\psi}=1$ its integral $  A=\oint dz \psi(z) $ is a nonlocal
operator of conformal dimension zero (invariant under conformal mapping $z\to w$ 
\begin{equation}
  \label{eq:50}
  A\to \oint dz \psi(z) \left(\frac{\partial w}{\partial z}\right)=\oint dw \psi(w)
\end{equation}
Then
\begin{multline}
  \label{eq:51}
  [L_{n},A]=\oint dz [L_{n},\psi(z)]=\\\oint dz  (n+1) z^{n}\psi(z)+z^{n+1}\partial \psi(z)=\\
  \oint dz \partial (z^{n+1}\psi (z))=0
\end{multline}
Screening operators are constructed from vertex operators
\begin{equation}
  \label{eq:52}
  V_{\pm}=V_{\alpha_{\pm}}, \quad \alpha_{\pm}=\alpha_{0}\pm \sqrt{\alpha_{0}^{2}+1}, \quad
  \alpha_{+}\alpha_{-}=-1,\quad \alpha_{+}+\alpha_{-}=2\alpha_{0}
\end{equation}
\begin{equation}
  \label{eq:53}
  Q_{\pm}=\oint dz V_{\pm}(z)=\oint dz e^{i\sqrt{2}\alpha_{\pm}\varphi(z)}
\end{equation}
Now for two-point function we can write
\begin{equation}
  \label{eq:54}
  \left<V_{\alpha}(z)V_{\alpha}(w)Q_{-}^{m}Q_{+}^{n}\right>,
\end{equation}
with neutrality condition $2\alpha+n\alpha_{+}+m\alpha_{-}=2\alpha_{0}$,  so $2\alpha$ is an integer
combination of $\alpha_{-}$ and $\alpha_{+}$.
So we obtain Kac formula for conformal weights of fields in minimal models
\begin{equation}
  \label{eq:55}
  \alpha_{r,s}=\frac{1}{2}(1-r)\alpha_{+}+\frac{1}{2}(1-s)\alpha_{-}\quad 
  h_{r,s}(c)=\frac{1}{4}(r\alpha_{+}+s\alpha_{-})^{2}-\alpha_{0}^{2}
\end{equation}

If we calculate three-point function and require operator algebra to be closed we can obtain
conditions on possible values of central charge or background charge $\alpha_{0}$.

Computation of three-point function relies on following trick: find four-point function as a
solution of differential equation following from null-vector condition and then recover structure
constants for three-point function from the limit $z\to 0$, where $z$ is cross-ratio.

Structure constants in three-point function are equal to fusion coefficients for corresponding
fields, for example substitute fusion rule \eqref{eq:39} to three-point function

  \begin{equation}
    \label{eq:56}
    \langle
    \varphi_{1}(x_{1})\varphi_{2}(x_{2})\varphi_{3}(x_{3})\rangle=\frac{C^{abc}_{123}}{x_{12}^{\Delta_{1}+\Delta_{2}-\Delta_{3}}
      x_{23}^{\Delta_{2}+\Delta_{3}-\Delta_{1}}x_{13}^{\Delta_{3}+\Delta_{1}-\Delta_{2}}}.     
  \end{equation}

Assume that we know four-point function
$\left<\varphi_{1}(z_{1})\varphi_{2}(z_{2})\varphi_{3}(z_{3})\varphi_{4}(z_{4})\right>$, then we
take limits $z_{1}\to z_{2}$ and $z_{3}\to z_{4}$ and obtain product of structure constants
$C_{12}^{r}C_{34}^{r}$ as a coefficient before $\frac{1}{z_{24}^{r}}$.

To compute four-point function we use conformal invariance to send $z_{1}\to 0$, $z_{2}\to z$
$z_{3}\to 1$ and $z_{4}\to \infty$:
\begin{equation}
  \label{eq:58}
  \left<\varphi_{1}(z_{1})\varphi_{2}(z_{2})\varphi_{3}(z_{3})\varphi_{4}(z_{4})\right>=\left<
    V_{r_{1},s_{1}}(z_{1}) V_{r_{2},s_{2}}(z_{2}) V_{r_{3},s_{3}}(z_{3}) V_{-r_{4},-s_{4}}(z_{4})
    Q_{+}^{r} Q_{-}^{s}\right>,
\end{equation}
where neutrality condition gives us
\begin{equation}
  \label{eq:59}
  \begin{array}{c}
    r_{4}=r_{1}+r_{2}+r_{3}-2r-2\\
    s_{4}=s_{1}+s_{2}+s_{3}-2s-2
  \end{array}
\end{equation}
then 
\begin{multline}
  \label{eq:58}
  \left<\varphi_{1}(0)\varphi_{2}(z)\varphi_{3}(1)\varphi_{4}(\infty)\right>=\left<
    V_{r_{1},s_{1}}(0) V_{r_{2},s_{2}}(z) V_{r_{3},s_{3}}(1) V_{-r_{4},-s_{4}}(\infty)
    Q_{+}^{r} Q_{-}^{s}\right>=\\
  z^{\frac{1}{3}\sum h_{i}-h_{1}-h_{2}}(1-z)^{\frac{1}{3}\sum h_{i}-h_{2}-h_{3}} G(z)
\end{multline}

The simplest case is when $\varphi_{2}=\varphi_{(2,1)}$, then we need to compute integral
\begin{multline}
  \label{eq:60}
  \left<
    V_{r_{1},s_{1}}(0) V_{2,1}(z) V_{r_{3},s_{3}}(1) V_{-r_{4},-s_{4}}(\infty)
    Q_{+}\right>=\\
  \oint dw \left<     V_{r_{1},s_{1}}(0) V_{2,1}(z) V_{r_{3},s_{3}}(1) V_{-r_{4},-s_{4}}(\infty)
    V_{+}(w)\right>=\\
  z^{2\alpha_{2,1}\alpha_{r_{1},s_{1}}}  (1-z)^{2\alpha_{2,1}\alpha_{r_{3},s_{3}}} G(z)
\end{multline}
Substitute \eqref{eq:41} and in holomorphic sector we obtain
\begin{equation}
  \label{eq:61}
  \oint dw w^{a}(w-1)^{b}(w-z)^{c},
\end{equation}
where
\begin{equation}
  \label{eq:62}
  a=2\alpha_{+}\alpha_{r_{1},s_{1}},\quad b=2\alpha_{+}\alpha_{r_{3},s_{3}}, c=2\alpha_{+}\alpha_{2,1}
\end{equation}

The difficulty is in choice of integration contour, we need to choose two independent contours which
will give us two independent solutions of differential equation \eqref{eq:33}.

Two choices of contour \cite{difrancesco1997cft} give integrals
\begin{equation}
  \label{eq:63}
  \begin{array}{ll}
    I_{1}(a,b,c; z) &= \int_{1}^{\infty} dw w^{a} (w-1)^{b} (w-z)^{c}\\
    I_{1}(a,b,c; z) &= \int_{0}^{z} dw w^{a} (1-w)^{b} (w-z)^{c}\\
    &=z^{1+a+b+c}\int_{0}^{1}dw  w^{a}(1-w)^{b}(1-zw)^{c}
  \end{array}
\end{equation}
These integrals can be calculated by power series expansion of terms with $z$, use of integral
formula for beta-function 
\begin{equation}
  \label{eq:64}
    B(x,y)=\frac{\Gamma(x)\Gamma(y)}{\Gamma(x+y)} = \int_{0}^{1} t^{x-1} (1-t)^{y-1} dt =
    \int_{0}^{\infty} t^{x-1} (1+t)^{-x-y} dt
  \end{equation}
then convert power series to hypergeometric function definition using $\Gamma(t+1)=t\Gamma(t)$
\begin{equation}  
  F(\lambda,\mu,\nu; z)=\sum_{k=0}^{\infty}\frac{1}{k!}\frac{(\lambda)_{k}(\mu)_{k}}{(\nu)_{k}}
  z^{k}, \;\mbox{where}\; (x)_{k}=x(x-1)\dots (x-k+1)
\end{equation}
Final result for conformal block is
\begin{eqnarray}
  \label{eq:65}
  I_{1}(a,b,c;z)=\frac{\Gamma(-a-b-c-1)\Gamma(b+1)}{\Gamma(-a-c)} F(-c,-a-b-c-1,-a-c; z)\\
  I_{2}(a,b,c;z)=z^{1+a+b+c}\frac{\Gamma(a+1)\Gamma(c+1)}{\Gamma(a+c+2)} F(-b,a+1,a+c+2; z)\\
  G(z,\bar z)=\sum_{i,j=1,2}X_{ij} I_{i}(z) \bar I_{j}(\bar z)
\end{eqnarray}
Calculation of coefficients $X_{ij}$ can be made using monodromy properties of correlation function
around singularities. It is equivalent to operator product expansion of primary fields. Functions
$I_{1}, I_{2}$ are in correspondence with intermediate fields in OPE for pairs of primary fields in
four-point function. Different ways to calculate using OPE correspond to different conformal blocks:
\begin{equation}
  \label{eq:66}
  \mathcal{G}(z,\bar z) =
  \left<\varphi_{1}(z_{1})\varphi_{2}(z_{2})\varphi_{3}(z_{3})\varphi_{4}(z_{4})\right> =
  \sum_{r,s;\bar r, \bar s}\mathcal{F}_{r,s}(z) \bar{\mathcal{F}}_{\bar r,\bar s}(\bar z)
\end{equation}
where $\mathcal{F}_{r,s}$ corresponds to a field $\varphi_{(r,s)}$ in OPE
\begin{equation}
  \label{eq:67}
  \varphi_{(r,s)}\in \varphi_{(r_{1},s_{1})}\times \varphi_{(r_{2},s_{2})}
\end{equation}

Complete results for four-point functions in minimal model with Coulomb gas representation are
presented in paper \cite{dotsenko1985four}.

\subsubsection{Liouville theory}
\label{sec:liouville-theory-1}

Liouville theory is very similar to Coulomb gas picture for minimal models. Scalar curvature at
infinity is $4\pi b$.

Stress-energy tensor is
\begin{equation}
  \label{eq:68}
  T(z)=-(\partial \varphi)^{2}+Q \partial^{2}\varphi
\end{equation}
Central charge is related to $Q$ by formula \eqref{eq:19}, similarly to minimal models
\eqref{eq:49}.

It can be shown (add reference here? \cite{nakayama2004liouville},\cite{teschner2001liouville}) that 
\begin{equation}
  \label{eq:34}
  V_{-\frac{b}{2}}\times V_{\alpha}= [V_{\alpha-\frac{b}{2}}]+[V_{\alpha+\frac{b}{2}}]
\end{equation}

We can introduce structure constants of this operator product expansion \cite{fateev2000boundary}
\begin{equation}
  \label{eq:35}
   V_{-\frac{b}{2}} V_{\alpha}= C_{+}[V_{\alpha-\frac{b}{2}}]+C_{-}[V_{\alpha+\frac{b}{2}}]
\end{equation}

This structure constants are equal to constants in three-point functions and can be computed by
screening integrals. For $C_{+}$ we have neutrality condition
$\alpha-\frac{b}{2}+Q-\alpha+\frac{b}{2}=Q$, so we do not need screening integrals and $C_{+}=1$.
For $C_{-}$ we screen by first order of Liouville interaction $-\mu\int e^{2b\varphi} d^{2}x$ 
(Add notes from Zamolodchikov lectures p.131,141 \cite{zamolodchikovlectures}): 
\begin{multline}
  \label{eq:36}
  C_{-}=-\mu\int d^{2}x \left<V_{\alpha}(0)V_{-\frac{b}{2}}(1) e^{2b\varphi(x)}
    V_{Q-\alpha-\frac{b}{2}}(\infty)\right>=\\
  \int d^{2}x x^{2\alpha b} (x-1)^{-b^{2}}=\\
  -\mu\frac{\pi\gamma(2b\alpha-1-b^{2})}{\gamma(-b^{2})\gamma(2b\alpha)},
\end{multline}
where
\begin{equation}
  \label{eq:37}
  \gamma(x)=\frac{\Gamma(x)}{\Gamma(1-x)}
\end{equation}

To calculate two-point function
$\left<V_{\alpha}(x)V_{\alpha}(0)\right>=\frac{D(\alpha)}{|x|^{2\Delta_{\alpha}}}$ we consider
auxiliary three-point function
\begin{equation}
  \label{eq:38}
  \left<V_{\alpha}(x_{1})V_{\alpha+\frac{b}{2}}(x_{2})V_{-\frac{b}{2}}(z)\right>
\end{equation}
Then we consider two operator product expansions when $z\to x_{1}$ and when $z\to x_{2}$. In the
first case only the term $C_{-} V_{\alpha+\frac{b}{2}}$ survives and we get
$C_{-}D\left(\alpha+\frac{b}{2}\right)$. In the second case we get $C_{+}D(\alpha)$. Since $C_{+}=1$
we get functional equation on $D$:
\begin{equation}
  \label{eq:57}
  \frac{D\left(\alpha+\frac{b}{2}\right)}{D(\alpha)}=\frac{1}{C_{-}(\alpha)}
\end{equation}
The solution of this equation is
\begin{equation}
  \label{eq:69}
  D(\alpha)=(\pi\mu\gamma(b^{2}))^{\frac{Q-2\alpha}{b}} \frac{\gamma(2b\alpha-b^{2})}{b^{2}\gamma\left(2-\frac{2\alpha}{b}+\frac{1}{b^{2}}\right)}
\end{equation}

Result for three-point function can be obtained in the same way by considering auxiliary four-point
function that satisfy second order differential equation due to degeneracy of the field $V_{-\frac{b}{2}}$.

\subsection{Boundary Liouville theory}
\label{sec:bound-liouv-theory}

Boundary Liouville theory:
\begin{equation}
  \label{eq:12}
  S=\oint_{S^{1}} \phi \mathcal{N} \phi |dw| + \oint_{S^{1}} \mu_{B} e^{b\phi} + \mu\int_{D} e^{2b\phi} dw d\bar{w}
\end{equation}

\begin{equation}
  \label{eq:11}
  G^{r}(\mu,\mu_{B}) = \mu \frac{\left<V_{\frac{b}{2}} B_{-\frac{1}{b}}\right>}{\left<V_{\frac{b}{2}}\right>}
\end{equation}

In paper \cite{fateev2000boundary} following expression was obtained for one-point boundary
correlation function
\begin{equation}
  \label{eq:22}
  \left<V_{\alpha}(x)\right>=\frac{U(\alpha|\mu_{B})}{|z-\bar z|^{2\Delta_{\alpha}}}
\end{equation}

\begin{equation}
  \label{eq:21}
  U(\alpha)=\frac{2}{b}\left(\pi\mu\gamma( b^{2})\right)^{\frac{Q-2\alpha}{2b}} \Gamma(2b\alpha-b^{2})
  \Gamma\left(\frac{2\alpha}{b}-\frac{1}{b^{2}}-1\right) \cosh (2\alpha-Q)\pi s
\end{equation}
where $s$ is determined from the equation
\begin{equation}
  \label{eq:23}
  \cosh^{2} \pi b s=\frac{\mu_{B}^{2}}{\mu}\sin \pi b^{2}
\end{equation}
and $\gamma(x)=\frac{\Gamma(x)}{\Gamma(1-x)}$.

This one point function is related to generating function of surfaces with a conic singularity of given length and area in
two-dimensional gravity theory. 
Partition function for  surfaces with a conic singularity of given length and area in
two-dimensional gravity theory is denoted by $Z_{\alpha}(A,l)$. Canonical transformation gives us
\begin{equation}
  \label{eq:24}
  W_{\alpha}(l,\mu)=\int_{0}^{\infty} \frac{dA}{A} e^{-\mu A}Z_{\alpha}(A,l)
\end{equation}

For the partition function of surfaces with given area $A$ and length $l$ and conical singularity
with angle $a(\alpha)$ following expression is obtained:
\begin{equation}
  \label{eq:29}
  Z_{\alpha}(A,l)=\frac{1}{b}\frac{\Gamma(2\alpha b
    -b^{2})}{\Gamma\left(1+\frac{1}{b^{2}}-\frac{2\alpha}{b}\right)} \left(\frac{l\Gamma(b^{2})}{2
      A}\right)^{\frac{(Q-2\alpha)}{b}} \exp\left(-\frac{l^{2}}{4 A \sin \pi b^{2}}\right)
\end{equation}

How is it connected with the equation \eqref{eq:21}?

Note that number of surfaces of given area $A$ and length $l$ is given by a canonical ensemble
(parameters $N$, $E$)
\begin{equation}
  \label{eq:25}
  N(l,A) = \sum_{\mbox{surfaces}} \delta(\mbox{Area}-A)\delta(\mbox{Length}-l)
\end{equation}
\begin{equation}
  \label{eq:26}
  F=\log Z
\end{equation}
Using Legendre transformation ($f^{*}(p)=\sup_{x\in I} (xp - f(x))$ $x\to \mbox{tangent}$) we move
to the description with bulk and boundary interaction constants. 
\begin{equation}
  \label{eq:27}
  \min_{\mu}(F(\mu)-\mu A) = S = \log N(A) \quad \mbox{Enthropy}
\end{equation}
Thus we obtain grand canonical ensemble (number of particles is not fixed, $w_{n}=\frac{1}{Z}
e^{\frac{-(E_{n}-\mu N)}{kT}}$)

$Z(\mu_{B},\mu) = \sum_{\mbox{surfaces}} e^{-\mu_{B}l -\mu A}=\int D\varphi e^{-S[\varphi]}$
\begin{equation}
  \label{eq:28}
  \int D\varphi = \int dA dl \int D\phi \quad \mbox{but} \quad \int D\phi = N(A,l)
\end{equation}
It is useful to note different definition of integration measure in \cite{fateev2000boundary}: $\int
\frac{d A}{A}\frac{d l }{l} N(A,l)$

\begin{multline}
  \label{eq:30}
  \left< 1 \right>_{\alpha}=\int \left< V_{\alpha}(z)\right> dz = \int dz \sum_{\mbox{surfaces with cone
    }\alpha\mbox{ at }z} e^{-S[\mbox{surface}]}=\\
  \int \frac{dl}{l} \int dz  \sum_{\mbox{surfaces with cone
    }\alpha\mbox{ at }z \mbox{ of length }l}e^{-S[\mbox{surface}]}=\\
  \int\frac{dl}{l} W_{a}(l) e^{-\mu_{B}l}
\end{multline}

\begin{multline}
  \label{eq:31}
  Z_{\alpha}(\mu,\mu_{B}) = \int \left< V_{\alpha}(z)\right> dz=\\\int D\varphi\int d^{2}z \exp\left(-\frac{1}{4\pi} \int d^{2}z
    b(\nabla\varphi)^{2}-\mu_{B}\oint e^{b\varphi}dx -\mu\int d^{2}z e^{2b\varphi}\right.\\
  \left.-\alpha\int
  d^{2}w \varphi(w) \delta(z-w)\right)
\end{multline}
So we see that one point boundary correlation function gives us partition function for surfaces of
given length and area.

\subsection{Computation of one-point function}
\label{sec:comp-one-point}

To actually compute bulk one-point function \eqref{eq:22} in boundary Liouville field theory  we
consider auxiliary bulk two-point function
\begin{equation}
  \label{eq:70}
  G_{\alpha,-\frac{b}{2}}=\left<V_{\alpha}(x)V_{-\frac{b}{2}}(z)\right>
\end{equation}
Since we consider theory on upper half-plane, this two-point function satisfy the same equation as
four-point function $\left<V_{\alpha}(x)V_{-\frac{b}{2}}(z)V_{\alpha}(\bar x)V_{-\frac{b}{2}}(\bar
  z)\right>$ in the whole plane. (Zero mode?) 
(Note: $\left<\phi(z,\bar z)\right>_{b}\sim\left<\phi(z)\phi(\bar z)\right>=\frac{1}{(z-\bar
  z)^{2h}}$, then $\left<\phi(y)\right>\sim \frac{1}{y^{2h}}$)

Since our four-point function satisfy second order ODE, it can be presented as a linear combination
of two independent solutions -- conformal blocks. Change of variables allows to rewrite the equation
as a hypergeometric equation, so conformal blocks are given by hypergeometric functions:
\begin{equation}
  \label{eq:71}
  \mathcal{G}_{\pm}(x,z)=\frac{|x-\bar x|^{2\Delta_{\alpha}-2\Delta_{-\frac{b}{2}}}}{|z-\bar
    x|^{4\Delta_{\alpha}}} \mathcal{F}_{\pm}(\eta),
\end{equation}
where $\Delta_{-\frac{b}{2}}=-\frac{1}{2}-\frac{3b^{2}}{4}$ and 
\begin{equation}
  \label{eq:72}
  \eta=\frac{(z-x)(\bar z-\bar x)}{(z-\bar x)(\bar z - x)}
\end{equation}
\begin{eqnarray}
  \label{eq:73}
  F_{+}(\eta)=\eta^{\alpha b}(1-\eta)^{-\frac{b^{2}}{2}} F(2\alpha b -1-2b^{2},-b^{2},2\alpha
  b-b^{2};\eta)\\
  F_{-}(\eta)=\eta^{1+b^{2}-\alpha b}(1-\eta)^{-\frac{b^{2}}{2}} F(-b^{2},1-2\alpha b,2+b^{2}-2\alpha
  b;\eta)
\end{eqnarray}
Coefficients before independent solutions $\mathcal{G}_{\pm}$ are fixed by substituting OPE
\eqref{eq:35} (substitute it to equation?):
\begin{equation}
  \label{eq:74}
  G_{\alpha,-\frac{b}{2}}=C_{+}(\alpha) U\left(\alpha-\frac{b}{2}\right) \mathcal{G}_{+}(x,z)+
  C_{-}(\alpha) U\left(\alpha+\frac{b}{2}\right) \mathcal{G}_{-}(x,z)
\end{equation}
Next we need to consider OPE when both operators $V_{\alpha}(x), V_{-\frac{b}{2}}(z)$ approach the
boundary.

Bulk operator $V_{-\frac{b}{2}}$ near the boundary gives rise to two boundary families $B_{0}$ and
$B_{-b}$. The fusion of $B_{0}=I$ to $V_{\alpha}$ is $\left<V_{\alpha}(x)
  B_{0}(0)\right>=\frac{R(\alpha,0|\mu_{B})}{|z-\bar z|^{2\Delta_{\alpha}}}$, and the fusion of
$V_{-\frac{b}{2}}$

\section{Quantum gravity from random graphs}
\label{sec:quantum-gravity-from}

There is another approach to quantum gravity. Instead of field theory on a surface one can consider
random discrete surfaces and statistical models on them. Random triangulations or quadrangulations
of a surface can be represented by dual random graphs \cite{kazakov1986ising}.  The matter is
introduced as a lattice statistical model on such random graph. Then computation of partition
function (or correlation functions) for matter on random surface is equivalent to summation over
random planar Feynman diagrams for certain matrix model.

Using this approach Ivan Kostov was able to obtain disk partition function
\cite{kostov2004boundary,kostov2003boundary}. His results are as follows.

\begin{equation}
  \label{eq:95}
  W(\tau) = -\frac{M^{g}}{2\sin \pi g} \cosh g\tau,
\end{equation}
where parameters $M$, $\tau$ are defined by
\begin{equation}
  \label{eq:96}
  \mu=\frac{1}{\pi} \frac{\Gamma\left(1-\frac{1}{g}\right)}{\Gamma\left(\frac{1}{g}\right)} M^{2}
\end{equation}
\begin{equation}
  \label{eq:97}
  \mu_{B}=\frac{\Gamma\left(1-\frac{1}{g}\right)}{\pi} M\cosh \tau
\end{equation}

Here $g=\frac{1}{b^{2}}$. The result \eqref{eq:95} is supposed to coincide with the results of
\cite{fateev2000boundary} in the following sense:
\begin{equation}
  \label{eq:98}
  \partial_{\mu} W = \partial_{\mu_{B}} U
\end{equation}

\section{Scaling limit and CFT on a cone}
\label{sec:scaling-limit-cft}

Here we argue that scaling limit is connected with correlation
functions of CFT on a cone. Then we introduce recurrence relations on
the angle of the cone. We show that this recurrent relation is
equivalent to fusion rules of degenerate field in CFT.

\section{CFT fusion coefficients and Bessel's functions}
\label{sec:cft-fusi-coeff}

In this section we solve recurrent relation using fusion rules and
show that scaling function is equal to Bessel's function.

\section{Conclusion}
\label{sec:conclusion}
Some concluding remarks are here.

\bibliography{bibliography}{} 
\bibliographystyle{utphys}

\end{document}